### 2020-02-04 analysis/step1_get_drawings/THU_Drawings/1_checksubjects/subID_check.Rmd
port experiments & analyses for photodraw from kiddraw to new repo

---
title: "tsinghua-subids"
author: "Bria Long"
date: "1/16/2020"
output: html_document
---

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

```{r}
is_test = c('Ipad5',
'P0000',
'Test',
'Test4',
'Test5','Ipad2_tes3','Ipad1','ipad4_Test1','ipad4_Test1')

# ids in database
IDs_DB <- read_csv('subject_logs/subIDs_afterNov15.csv') %>%
  rowwise() %>%
  mutate(is_test = str_detect(subID,'test')) %>%
  filter(is_test==FALSE) %>%
  mutate(ID_Database = str_replace_all(subID,' ','')) %>%
  mutate(ID_Database = str_replace_all(ID_Database,'-','_'))  %>%
  mutate(ID_Database = str_replace_all(ID_Database,'__','_')) %>%
  mutate(ID_Database = str_replace_all(ID_Database,'I','i'))  %>%
  mutate(ID_Database = str_to_upper(ID_Database))  %>%
  mutate(iPad_Used = str_split(ID_Database,"[_]")[[1]][1],Short_SubID = str_split(ID_Database,"[_]")[[1]][2])

# ids in subject log
IDs_SL <- read_csv('subject_logs/KIDDRW_THU_cleaned.csv') %>%
  select(age, gender, ID_age, CB, ID, note,db_mismatch) %>%
  rowwise() %>%
  mutate(ID_SubjectLog = str_replace_all(ID,' ','')) %>%
  mutate(ID_SubjectLog = str_replace_all(ID_SubjectLog,'-','_'))  %>%
  mutate(ID_SubjectLog = str_replace_all(ID_SubjectLog,'__','_')) %>%
  mutate(ID_SubjectLog = str_replace_all(ID_SubjectLog,'I','i'))  %>%
  mutate(ID_SubjectLog = str_to_upper(ID_SubjectLog))  %>%
  mutate(iPad_Used = str_split(ID_SubjectLog,"[_]")[[1]][1],Short_SubID = str_split(ID_SubjectLog,"[_]")[[1]][2])

merged <- IDs_SL %>%
  left_join(IDs_DB, by=c("iPad_Used","Short_SubID"))

mismatch <- merged %>%
  filter(is.na(ID_Database)) %>%
  select(-db_mismatch, -subID, -is_test, -note)

db_only <- IDs_DB %>%
  left_join(IDs_SL, by=c("iPad_Used","Short_SubID")) %>%
  filter(is.na(ID_SubjectLog))

write_csv(db_only,'db_only_IDs.csv')
write_csv(mismatch,'no_db_entry.csv')

```

```{r}

# no_ipad_assigned = c("IPA2_THU9F41"  , "IPAD_THU5F6"  , "IPAD_THU6F13" ,  "IPAD_THU7M12"  , "IPAD_THU8M21" )
# 
# #mismatch candidates
# 
# mismatch_maybes <- c('IPAD1_1THU6F30',

match_count=0
subid_count=0
sorted_DB_Ids=array()
reported_IDS <- unique(reported_IDS)
db_IDS <- IDs$ID_Database
for (this_subid in reported_IDS){
  subid_count = subid_count +1
  match = !is.na(str_detect(this_subid, db_IDS))

  # if any matches
  if (sum(match)>0){
  match_count = match_count +1
  match_ind = which(!is.na(str_locate(this_subid, db_IDS)))[1]

  sorted_DB_Ids[subid_count] = db_IDS[match_ind]
  }
  else{
  sorted_DB_Ids[subid_count] = 'NA'
  }

}
```

```{r}
# ID_matches = bind_cols(data.frame(reported_IDS), data.frame(sorted_DB_Ids))
# num_matches = sum(!is.na(ID_matches$sorted_DB_Ids))
# need_matches = ID_matches$reported_IDS[is.na(ID_matches$sorted_DB_Ids)]

```

```{r}
db_sorted = sort(IDs$ID_Database)
not_matched_db <- as_tibble(db_sorted) %>%
  filter(!is.na(value)) %>%
  filter(!value %in% sorted_DB_Ids) %>%
  rowwise() %>%
  mutate(iPad_Used = str_split(value,"[_]")[[1]][1],SubID = str_split(value,"[_]")[[1]][2]) %>%
  mutate(age = str_split_fixed(SubID,'THU',2)[,2])

need_matches <- as_tibble(need_matches) %>%
  rowwise() %>%
  mutate(iPad_Used = str_split(value,"[_]")[[1]][1],SubID = str_split(value,"[_]")[[1]][2]) %>%
  mutate(maybe_match = SubID %in% not_matched_db$SubID)

super_confused <- need_matches %>%
  filter(maybe_match==FALSE)

## typo
IPAD1_1THU6F30 = IPAD1_THU6F30

# 
# write_csv(not_matched_db, "non-matched-db.csv")
# write_csv(as_tibble(need_matches), "Tsinghua_DB_need_matches.csv")
# db_sorted = sort(IDs$ID_Database)

```