### 2021-04-28 analysis/step1_get_drawings/THU_Drawings/2_subject_log_checks/Step1_subid_clean_and_check.Rmd
scripts to automatically merge subject logs and resave darwings

---
title: "tsinghua-subids"
author: "Bria Long"
date: "1/16/2020"
output: html_document
---

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
library(here)
theme_set(theme_few())
```

## Read in manually matched IDs
```{r}
IDs_match <- read_csv('subject_logs/correct_matches.csv') %>%
  rename(ID_SubjectLog = ID_SubjectLog_cleaned) %>%
  filter(!is.na(ID_Database_raw)) %>% # 3 subjects for whom we don't have matches at all
  mutate(ID_Database = str_replace_all(ID_Database_raw,' ','')) %>%
  mutate(ID_Database = str_replace_all(ID_Database,'-','_'))  %>%
  mutate(ID_Database = str_replace_all(ID_Database,'__','_')) %>%
  mutate(ID_Database = str_replace_all(ID_Database,'I','i'))  %>%
  mutate(ID_Database = str_to_upper(ID_Database))  %>%
  rowwise() %>%
  mutate(iPad_Used = str_split(ID_Database,"[_]")[[1]][1],Short_SubID = str_split(ID_Database,"[_]")[[1]][2])

```

## Read in Ids from  database

```{r}
### test ids
is_test = c('Ipad5',
'P0000',
'Test',
'Test4',
'Test5','Ipad2_tes3','Ipad1','ipad4_Test1','ipad4_Test1')
```

```{r}
IDs_DB <- read_csv('subject_logs/subIDs_afterNov15.csv') %>%
  rowwise() %>%
  mutate(is_test = str_detect(subID,'test')) %>%
  filter(is_test==FALSE) %>%
  mutate(ID_Database = str_replace_all(subID,' ','')) %>%
  mutate(ID_Database = str_replace_all(ID_Database,'-','_'))  %>%
  mutate(ID_Database = str_replace_all(ID_Database,'__','_')) %>%
  mutate(ID_Database = str_replace_all(ID_Database,'I','i'))  %>%
  mutate(ID_Database = str_to_upper(ID_Database))  %>%
  mutate(iPad_Used = str_split(ID_Database,"[_]")[[1]][1],Short_SubID = str_split(ID_Database,"[_]")[[1]][2]) %>%
  rename(ID_Database_raw = subID)
```

### Read in IDs from subject log
```{r}
# ids in subject log
IDs_SL <- read_csv('subject_logs/subject_log_sept2020.csv') %>%
  select(Age, Sex, ID, CB, note,Date) %>%
  rowwise() %>%
  mutate(ID_SubjectLog = str_replace_all(ID,' ','')) %>%
  mutate(ID_SubjectLog = str_replace_all(ID_SubjectLog,'-','_'))  %>%
  mutate(ID_SubjectLog = str_replace_all(ID_SubjectLog,'__','_')) %>%
  mutate(ID_SubjectLog = str_replace_all(ID_SubjectLog,'I','i'))  %>%
  mutate(ID_SubjectLog = str_to_upper(ID_SubjectLog))  %>%
  mutate(iPad_Used = str_split(ID_SubjectLog,"[_]")[[1]][1],Short_SubID = str_split(ID_SubjectLog,"[_]")[[1]][2]) 
```

```{r}

# need to use specdal spreadsheet for these subjects
to_fix <- IDs_SL %>% # get ids from subject log
  filter(ID_SubjectLog %in% IDs_match$ID_SubjectLog) %>% # get set of subjects in subject log that we had to manually match 
  left_join(IDs_match %>% select(ID_SubjectLog, ID_Database_raw, ID_Database))  
```
```{r}
# can merge with subject log here
merged <- IDs_SL %>%
  filter(!ID_SubjectLog %in% IDs_match$ID_SubjectLog) %>%
  left_join(IDs_DB, by=c("iPad_Used","Short_SubID")) %>%
  full_join(to_fix)

## assumption is didn't record these due to technical eerrors
mismatch <- merged %>%
  filter(is.na(ID_Database)) %>%
  select(-is_test, -note)

write_csv(mismatch,'no_db_entry_sept2020_tech_errors.csv')
```

## Save out information
```{r}
final_subject_info <-  merged %>%
  filter(!is.na(ID_Database))

write_csv(final_subject_info,here::here('data/THU/demographics/final_subject_info.csv'))
```

```{r}
subs_by_age <- final_subject_info %>%
  mutate(age_group = floor(Age)) %>%
  group_by(age_group) %>%
  summarize(count_subs = length(unique(ID_Database)))
```