### 2021-04-28 analysis/step4_main_analyses_both_sites/main_analysis.Rmd
remove old files

---
title: "Examine_classifications"
author: "Bria Long"
date: "4/9/2021"
output: html_document
---

```{r, libraries}
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(langcog)
library(forcats)
library(egg)
library(gridExtra)
library(reticulate)
library(readr)
library(ggplot2)
library(reshape2)
library(lme4)
library(stringr)
library(viridis)
library(MuMIn)
library(here)
theme_set(theme_few())
```

## Load compiled classification data
```{r load-classifications}
classification_data <- read.csv('compiled_classifications/Classification_Outputs2832.csv') %>%
  as.tibble() %>%
  # mutate(session_id = paste('cdm_',session_id,sep="")) %>%
  mutate(age_numeric = age) %>%
  mutate(age = paste('age',age,sep="")) %>%
  mutate(age = as.factor(age)) %>%
  mutate(category = target_label) %>% 
  # mutate(image_name = paste(category,'_sketch_', age,'_', session_id,'.png',sep="")) %>%
  select(-X) %>%
  mutate(site = case_when( is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "THU", 
                           !is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "CDM")) %>%
  mutate(site = as.factor(site))
```

## Need to rewrite some code to get meta-data for couchdb/mongodb dumps
```{r}
## Load in meta data from mongo-db database dumps
# metadata_photodraw <- read.csv('mongodb-output/Photodraw2_AllDescriptives_1475_images_final_CDM_photodraw_e2_no_blanks.csv') %>%
#   as.tibble() %>%
#   mutate(session_id = str_split_fixed(session_id,"_",2)[,2]) %>%
#   mutate(session_id = as.factor(session_id)) %>%
#   mutate(draw_duration_new = replace(draw_duration_new, draw_duration_new > 30,30))
# 
# ## join with classification data
# d <- classification_data %>%
#   left_join(metadata_photodraw) %>%
#   mutate(age_numeric = as.numeric(str_split_fixed(age,'age',2)[,2]))
```

## Read in meta data fed to classifications
```{r}
meta <- read_csv(here('data/compiled/meta/METADATA_kid.csv'))
```

## Compile some sanity checks on counts of draiwngs per site/condition/etc
```{r}
sanity_counts <- classification_data %>%
  left_join(meta, by=c('session_id' = 'session', 'category' = 'label')) %>%
  filter(age_numeric < 10) %>%
  group_by(condition,age_numeric, site) %>%
  summarize(num_drawings = n()) %>%
  kable()
```

```{r}
participant_counts <- d %>%
  group_by(age_numeric, site) %>%
  summarize(num_participants = length(unique(session_id)))

count_by_site <- d %>%
  group_by(site) %>%
  summarize(num_participants = length(unique(session_id)))

count_by_category <- d %>%
  group_by(category) %>%
  summarize(num_drawings = length(unique(session_id)), what = length(unique(index)))

what <- d %>%
  group_by(index) %>%
  summarize(num_ids = length(session_id)) %>%
  filter(num_ids >1)

count_by_id <- d %>%
  group_by(session_id) %>%
  summarize(num_drawings = length(unique(category))) 
```

265 participants were recruited from the San Jose Children’s Discovery museum, the Palo Alto Junior Museum & Zoo, and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. We aimed to recruit approximately 120 children aged 4-9 years of age after exclusions (i.e. 20 4-year-olds, 20 5-year-olds). In the US-based sample, 135 children participated; 6 participants were excluded, (3) for skipping more than 6 trials, and (3) for scribbling three or more times in a row; (6) participants were tested but their data was not recorded due to a technical error, leading to a final sample of `r count_by_site$num_participants[1]` children. In the China based sample, 140 children participated; (10) participants were tested but their data was not recorded due to a technical error with the remote database, leading to a final sample of `r count_by_site$num_participants[2]`. A complete breakdown of the number of subjects in each age group and site can be found in the Appendix, Table 1A.

Blank drawings were excluded from analysis, and drawings were randomly undersampled to have an equal number of drawings for each leave-one-out-classification. On average, each child contributed `r mean(count_by_id$num_drawings)` to analysis (min `r min(count_by_id$num_drawings)`, max = `r max(count_by_id$num_drawings)`) with leaving 232-234 drawings per category.

# Descriptive anlayses
```{r descriptives-across-age}
### How do our covariates change with age? Compute means and CIs; Group by age/category
## first join  data  
d <- classification_data %>% 
  left_join(meta, by=c('session_id' = 'session', 'category' = 'label'))

```

```{r}
cor_by_age <- d %>%  
  group_by(session_id,category,condition,age_numeric, site) %>%
  summarize(avg_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_cor") 
```

```{r}
base_size_chosen=12; smooth_alpha=.2
ggplot(cor_by_age, aes(age_numeric,mean, col = condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) + 
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  ylim(0,1) + 
  geom_smooth(span=10, alpha=.2) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_grid(~site)
```

### Correlation of item effects across sites
```{r}
item_effects <- d %>%
  group_by(session_id,age_numeric,category, site) %>%
  summarize(avg_sub_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects, aes(x=CDM, y=THU, color=age_numeric)) +
  # geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.5)) + 
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  # labs(x='Age', y='Classification accuracy') +
  # scale_color_viridis(option="D") + 
  # theme(legend.position = "none") + 
  # geom_smooth(col='grey',span=10, alpha=smooth_alpha) +
  ylim(0,1) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_wrap(~category)
```

```{r}
item_effects_no_age <- d %>%
  group_by(session_id,category, site) %>%
  summarize(avg_sub_cor = mean(correct_or_not)) %>%
  group_by(category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects_no_age, aes(x=CDM, y=THU, color=category)) +
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  geom_abline(intercept = 0, slope=1)
```

# Inferential stats
```{r}
full_model <- glmer(correct_or_not ~ condition*age_numeric+ site +
                        (1|session_id) +
                        (1|category),
      data = d, family="binomial")

summary(full_model)

```

# Witheffort covariates (not ready yet)
```{r}
draw_duration <- d %>%
  group_by(session_id,condition,age_numeric) %>%
  summarize(avg_draw_duration = mean(draw_duration_new)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_draw_duration")

num_strokes <- d %>%
  group_by(session_id,condition,age_numeric) %>%
  summarize(avg_num_strokes = mean(num_strokes)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_num_strokes")

avg_intensity <- d %>%
  group_by(session_id,condition,age_numeric) %>%
  summarize(avg_intensity = mean(mean_intensity)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_intensity")

```

```{r plot-descriptives-across-age}
## Make compiled plot of descriptives
base_size_chosen=18 # size of text in plots
smooth_alpha=.2

cor_by_age_plot_A = ggplot(cor_by_age, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) + 
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  # scale_color_viridis(option="D") + 
  theme(legend.position = "none") +
  # geom_smooth(col='grey',span=10, alpha=smooth_alpha) +
  # ggtitle('A') + 
  ylim(0,.75) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey")

p1=ggplot(draw_duration, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Draw duration (s)') +
  # scale_color_viridis(option="D") +
  theme(legend.position = "none") + 
  ylim(0,20) 
  # geom_smooth(col='grey', span = 10) +
  # ggtitle('B')

p2=ggplot(avg_intensity, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Ink used (mean intensity)') +
  # scale_color_viridis(option="D") +
  theme(legend.position = "none") + 
  ylim(.02,.08) 
  # geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
  # ggtitle('C')

p3=ggplot(num_strokes, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Number of strokes') +
  # scale_color_viridis(option="D") +
  theme(legend.position = "right") +
  ylim(0,15) 
  # geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
  # ggtitle('D')

# p4=ggplot(tracing_scores, aes(age_numeric,mean, color=age_numeric)) +
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
#   theme_few(base_size = base_size_chosen) +
#   labs(x='Age', y='Normalized tracing score') +
#   # scale_color_viridis(option="D") +
#   theme(legend.position = "none") + 
#   geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
#   ggtitle('E')
```