### 2021-04-28 analysis/step2_clean_metadata/save_batches_for_recognition.Rmd

---
title: "Photodraw2-Descriptives"
author: "Bria Long"
date: "2/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(assertthat)
library(here)
```

## Do you really want to rerun this?
```{r}
## Generally you don't -- this does a lot of copying and filtering based on the dataset preprocessing stage...
rerun=FALSE
```
## Import metadata
```{r}
all_meta <- read.csv(file = here::here('data/compiled/metadata/final_merged_metadata.csv')) %>%
  as_tibble() 
```

## looked at differences between full set of rendered drawings and cleaned metadata; removed excluded drawings
```{r}

if (rerun=TRUE){
  compiled_dataset_path = here::here('data/compiled/drawings/object_drawings')

  all_rendered_drawings = list.files(path=compiled_dataset_path, pattern = ".png$", recursive = TRUE) %>%
      as_tibble()  %>%
      rename(fullPath = value)

  prac_trials = c('this square','this shape','triangle','rectangle')
  object_drawings <- all_meta %>%
    filter(!category %in% prac_trials)

  # either duplicate sessions or no stroke data, see merge_and_clean_datasets.
  drawings_excluded <- all_drawings %>%
    filter(!fullPath %in% object_drawings$filename)

  ## DO ONCE
  file.remove(fs::path(compiled_dataset_path, drawings_excluded$fullPath))

  # should be none
  drawings_not_rendered <- object_drawings %>%
    filter(!filename %in% all_drawings$fullPath)

  assert_that(length(drawings_not_rendered$subID)==0)
  length(object_drawings$filename)
  length(all_drawings$fullPath)
}
```

# copy tracings from included THU children
```{r}
if (rerun==TRUE){
  compiled_tracings_path = here::here('data/compiled/drawings/tracings')
  tracings_THU = here::here('data/THU/drawings/tracings/')

  tracings_to_include_THU <- all_meta %>%
    filter(category %in% c('this square','this shape')) %>%
    filter(site == "THU")

  file.copy(fs::path(tracings_THU, tracings_to_include_THU$filename), fs::path(compiled_tracings_path, tracings_to_include_THU$filename))
}
```

# copy tracings from included CDM children
```{r}
if (rerun==TRUE){
compiled_tracings_path = here::here('data/compiled/drawings/tracings')
tracings_CDM = here::here('data/CDM/drawings/tracings/')

tracings_to_include_CDM <- all_meta %>%
  filter(category %in% c('this square','this shape')) %>%
  filter(site == "CDM")

file.copy(fs::path(tracings_CDM, tracings_to_include_CDM$filename), fs::path(compiled_tracings_path, tracings_to_include_CDM$filename))
}
```

# Randomly sample from groups of categories / ages / sites into subsets for different participants
```{r}
if (rerun==TRUE){
sample_set <- object_drawings
##
max_iter=21 
for (iteration in 1:max_iter) {
    if (iteration==max_iter){
      # whatever we have left if on last iteration
      print('last iteration...')
      print (length(sample_set$filename))
      write_csv(sample_set,here::here('data/compiled/recognition_rating_chunks/',paste0('chunk_',iteration,'_compiled_dataset.csv')))
    }
    # for all the other iterations
    else if (iteration<max_iter){

    this_chunk <- sample_set %>%
      group_by(age_numeric,category, site) %>%
      slice_sample(n=1, replace=FALSE)

    assert_that(length(unique(this_chunk$category))==12)
    assert_that(length(unique(this_chunk$age_numeric))==6)
    assert_that(length(unique(this_chunk$site))==2)

    print (length(this_chunk$filename))
    write_csv(this_chunk,here::here('data/compiled/recognition_rating_chunks/',paste0('chunk_',iteration,'_compiled_dataset.csv')))

    sample_set <- sample_set %>%
      filter(!filename %in% this_chunk$filename)

    # print (length(sample_set$filename))
    }
}
}

```