### 2021-04-28 analysis/step4_main_analyses/examine_photodraw_classifications.Rmd
move starter analysis scripts to new folder

---
title: "Examine_photodraw_classifications"
author: "Bria Long"
date: "4/2021"
output: html_document
---

```{r, libraries}
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(langcog)
library(forcats)
library(egg)
library(gridExtra)
library(reticulate)
library(readr)
library(ggplot2)
library(reshape2)
library(lme4)
library(stringr)
library(viridis)
library(MuMIn)
library(here)
theme_set(theme_few())
```

## Load compiled classification data
```{r load-classifications}
classification_data <- read.csv(here::here('data/compiled/compiled_classifications/Classification_Outputs2832.csv')) %>%
  as_tibble() %>%
  # mutate(session_id = paste('cdm_',session_id,sep="")) %>%
  mutate(age_numeric = age) %>%
  mutate(age = paste('age',age,sep="")) %>%
  mutate(age = as.factor(age)) %>%
  mutate(category = target_label) %>% 
  # mutate(image_name = paste(category,'_sketch_', age,'_', session_id,'.png',sep="")) %>%
  select(-X, -X.1, -index) %>%
  mutate(site = case_when( is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "THU", 
                           !is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "CDM")) %>%
  mutate(site = as.factor(site))
```

## Get drawing-specific metadata and join with main dataset
```{r}
## Load in meta data from mongo-db database dumps
CDM_metadata <- read.csv(here::here('data/CDM/drawing_metadata/Photodraw2_CDM_AllDescriptives_1922_images_final_CDM_photodraw_e2.csv')) %>%
  as_tibble() %>%
  mutate(session_id = str_split_fixed(session_id,"_",2)[,2]) %>%
  mutate(session_id = as.factor(session_id)) %>%
  mutate(draw_duration = replace(draw_duration_new, draw_duration_new > 30,30)) %>%
  select(session_id, condition, category, num_strokes, draw_duration, mean_intensity)

```

# for THU
```{r}
THU_final_subject_info = read_csv(here::here('data/THU/demographics/final_subject_info.csv')) %>%
  as_tibble() %>%
  select(ID_Database_raw, ID_Database) %>%
  rowwise() %>%
  mutate(ID_Database_raw = str_replace_all(ID_Database_raw, fixed(" "), "")) %>%
  mutate(ID_Database_raw = as.factor(ID_Database_raw)) %>%
  rename(session_id = ID_Database) #session_id = cleaned ID_Database anme for merging with CDM data

all_THU_metadata<- read.csv(here::here('data/THU/drawing_metadata/THU_Drawings_AllDescriptives_april2021_all.csv')) %>%
  select(subid, condition, category, num_strokes, draw_duration, mean_intensity, filename) %>%
  rowwise() %>%
  mutate(ID_Database_raw = str_replace_all(subid, fixed(" "), "")) %>% #subid = ID_Database_raw in this spreadsheet, I know, sorry
  mutate(ID_Database_raw = as.factor(ID_Database_raw))

# joint drawing metadata with subject info
THU_metadata <-  THU_final_subject_info %>%
  left_join(all_THU_metadata, by=c("ID_Database_raw")) %>%
  select(-ID_Database_raw, -subid, - filename)

# check nothing went wrong with merging
assert_that(sum(is.na(THU_metadata$num_strokes))==0)

```

### join both sets of metadata together
```{r}
all_meta <- CDM_metadata %>%
  full_join(THU_metadata)
```

### join with classification data -- which has age embedded
```{r}
d <- classification_data %>%
  left_join(all_meta, by=c('category','session_id'))  %>%
  mutate(age_numeric = as.numeric(str_split_fixed(age,'age',2)[,2])) %>%
  mutate(age_numeric = replace(age_numeric, age_numeric==10,9)) # two THU participants who are just barely 10-years old, including in 9-year-old group

# subids in d should be a superset of those in classification_data
check_ids <- classification_data %>%
  filter(!session_id %in% unique(all_meta$session_id))
assert_that(length(check_ids$age)==0) # all match up

```

### get rid of tech error drawings (see note)
```{r}
# bizarrely, there are 24 drawing for which stroke data didn't save -- technical error.
# this was checked directly in mongodb database -- no idea why, assuming bad internet.
tech_error_drawings <- d %>%
  filter(is.na(num_strokes))
# filter out terch error drawings from dataset (24)
d <- d %>%
  filter(!is.na(num_strokes)) 
```

```{r}
# only did one practice trial, not in subject log or classification data because no category level trials
missing_id <- CDM_metadata %>%
  filter(!session_id %in% d$session_id)
```

## Compile some sanity checks on counts of draiwngs per site/condition/etc
```{r}
sanity_counts <- d %>%
  group_by(condition,age_numeric, site) %>%
  summarize(num_drawings = n()) %>%
  kable()
```

```{r}
participant_counts <- d %>%
  group_by(age_numeric, site) %>%
  summarize(num_participants = length(unique(session_id)))

count_by_site <- d %>%
  group_by(site) %>%
  summarize(num_participants = length(unique(session_id)))

count_by_category <- d %>%
  group_by(category) %>%
  summarize(num_drawings = length(unique(session_id)))

count_by_id <- d %>%
  group_by(session_id) %>%
  summarize(num_drawings = length(unique(category))) 
```

265 participants were recruited from the San Jose Children’s Discovery museum, the Palo Alto Junior Museum and Zoo, and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. We aimed to recruit approximately 120 children between 4-9 years of age after exclusions (i.e. 20 4-year-olds, 20 5-year-olds, etc.). In the US-based sample, 135 children participated; 6 participants were excluded, (3) for skipping more than 6 trials, and (3) for scribbling three or more times in a row; (6) participants were tested but their data was not recorded due to a technical error, and (2) participants never made it past the practice trials, leading to a final sample of 121 children. In the China based sample, `r count_by_site$num_participants[2]` children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. A complete breakdown of the number of subjects in each age group and site can be found in the Appendix, Table 1A.

Blank drawings were excluded from analysis, and drawings were randomly undersampled to have an equal number of drawings for each leave-one-out-classification. On average, each child contributed `r mean(count_by_id$num_drawings)` to analysis (min `r min(count_by_id$num_drawings)`, max = `r max(count_by_id$num_drawings)`) with leaving 232-234 drawings per category.

# Descriptive anlayses
```{r descriptives-across-age}
### How do our covariates change with age? Compute means and CIs; Group by age/category
cor_by_age <- d %>%  
  group_by(session_id,category,condition,age_numeric, site) %>%
  summarize(avg_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_cor") 
```

```{r}
base_size_chosen=12; smooth_alpha=.2
ggplot(cor_by_age, aes(age_numeric,mean, col = condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) + 
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  ylim(0,1) + 
  geom_smooth(span=10, alpha=.2) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_grid(~site)
```

### Correlation of item effects across sites
```{r}
item_effects <- d %>%
  group_by(session_id,age_numeric,category, site) %>%
  summarize(avg_sub_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects, aes(x=CDM, y=THU, color=age_numeric)) +
  # geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.5)) + 
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  # labs(x='Age', y='Classification accuracy') +
  # scale_color_viridis(option="D") + 
  # theme(legend.position = "none") + 
  # geom_smooth(col='grey',span=10, alpha=smooth_alpha) +
  ylim(0,1) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_wrap(~category)
```

```{r}
item_effects_no_age <- d %>%
  group_by(session_id,category, site) %>%
  summarize(avg_sub_cor = mean(correct_or_not)) %>%
  group_by(category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects_no_age, aes(x=CDM, y=THU, color=category)) +
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  geom_abline(intercept = 0, slope=1)
```

# Inferential stats
```{r}
full_model <- glmer(correct_or_not ~ condition*age_numeric+ site +
                        (1|session_id) +
                        (1|category),
      data = d, family="binomial")

summary(full_model)

```

# Witheffort covariates (not ready yet)
```{r}
draw_duration <- d %>%
  group_by(session_id,condition,age_numeric) %>%
  summarize(avg_draw_duration = mean(draw_duration_new)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_draw_duration")

num_strokes <- d %>%
  group_by(session_id,condition,age_numeric) %>%
  summarize(avg_num_strokes = mean(num_strokes)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_num_strokes")

avg_intensity <- d %>%
  group_by(session_id,condition,age_numeric) %>%
  summarize(avg_intensity = mean(mean_intensity)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_intensity")

```

```{r plot-descriptives-across-age}
## Make compiled plot of descriptives
base_size_chosen=18 # size of text in plots
smooth_alpha=.2

cor_by_age_plot_A = ggplot(cor_by_age, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) + 
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  # scale_color_viridis(option="D") + 
  theme(legend.position = "none") +
  # geom_smooth(col='grey',span=10, alpha=smooth_alpha) +
  # ggtitle('A') + 
  ylim(0,.75) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey")

p1=ggplot(draw_duration, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Draw duration (s)') +
  # scale_color_viridis(option="D") +
  theme(legend.position = "none") + 
  ylim(0,20) 
  # geom_smooth(col='grey', span = 10) +
  # ggtitle('B')

p2=ggplot(avg_intensity, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Ink used (mean intensity)') +
  # scale_color_viridis(option="D") +
  theme(legend.position = "none") + 
  ylim(.02,.08) 
  # geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
  # ggtitle('C')

p3=ggplot(num_strokes, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Number of strokes') +
  # scale_color_viridis(option="D") +
  theme(legend.position = "right") +
  ylim(0,15) 
  # geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
  # ggtitle('D')

# p4=ggplot(tracing_scores, aes(age_numeric,mean, color=age_numeric)) +
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
#   theme_few(base_size = base_size_chosen) +
#   labs(x='Age', y='Normalized tracing score') +
#   # scale_color_viridis(option="D") +
#   theme(legend.position = "none") + 
#   geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
#   ggtitle('E')
```