### 2021-06-14 analysis/step5_tracing_evaluation/step2_evaluate_tracings/model_pred.Rmd
add tracing rendering and analysis scripts, update gitignore

---
title: "Model Prediction"
output: html_document
---

```{r}
library(tidyverse)
library(lme4)
library(groupdata2)
library(dplyr)
library(knitr) 
library(broom) 
library(hydroGOF)
library(MASS)
library(DescTools)
library(broom)
library(brms)
library(reshape2)
library(caret)
```

```{r}
load("vm_model.rda")
```

```{r}
# read the orig files
data_dir = here::here('data')
# filename = 'kiddraw_tracing_cdm_run_v4.csv'

files = dir(data_dir)
```

```{r}

count_files = 0
for (f in files) {
  count_files = count_files + 1 
  new_data <- read.csv(fs::path(data_dir,f))
      # select(-X)

  if (count_files == 1) {
    all_data <- new_data  
  }

  if (count_files > 1) {
  all_data <- all_data %>%
    full_join(new_data) 
  }
}
```

```{r}
length(unique(all_data$session_id))

```

```{r}
# predict on new data
ord.pred = predict(ord_wid, all_data, re_formula = NA)
ord.pred = data.frame(ord.pred)
all_data$rating = round(ord.pred$P.Y...1*1 + ord.pred$P.Y...2*2 + ord.pred$P.Y...3*3 + ord.pred$P.Y...4*4)
write.csv(all_data, paste0("rated_", "all_museumstation_filtered.csv"))
```