### 2021-06-14 analysis/step4b_download_recognition_data/preprocess_recognition_data.Rmd
scripts to preprocesss recognition data

---
title: "preprocess_human_recognition"
author: "Bria Long"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# compile all output csvs
```{r}
all_data=as_tibble()
for (batch in range(1,2)){
  this_batch = read.csv(paste0("csv_out/Devphotodraw_recognition_batch",batch,".csv")) %>% as_tibble() 
  all_data <- rbind(all_data, this_batch)
}

```

# preprocess image name / age
```{r}
d <- all_data %>%
  mutate(image_name_short = str_split_fixed(imageName,'object_drawings/',2)[,2]) %>%
  mutate(age_temp = str_split_fixed(image_name_short,'age',2)[,2]) %>%
  mutate(age = as.numeric(str_split_fixed(age_temp,'_',2)[,1]))
```

# filter out participants who don't finish
```{r}
included_subs <- d %>%
  group_by(subID) %>%
  summarize(num_drawings = length(unique(image_name_short))) %>%
  filter(num_drawings > 140)

used_all_options <- d %>%
  group_by(subID) %>%
  summarize(num_labels = length(unique(guessed_category)))

d <- d %>%
  filter(subID %in% included_subs$subID)
```

# quick check by age
```{r}
check_by_age <- d %>%
  group_by(age) %>%
  summarize(mean(guessed_category == drawn_category))
```