### 2021-06-14 analysis/step5_tracing_evaluation/step2_evaluate_tracings/model_pred.Rmd
update .rmd to make predictions on devphotodrawdata

---
title: "Use previously fit model to predict tracing scores"
output: html_document
---

```{r}
library(tidyverse)
library(lme4)
library(groupdata2)
library(dplyr)
library(knitr) 
library(broom) 
library(hydroGOF)
library(MASS)
library(DescTools)
library(broom)
library(brms)
library(reshape2)
library(caret)
library(here)
```

# Load model
```{r}
load("vm_model.rda")
```

# Read in tracing scores from airlab
```{r}
# read the orig files
all_data = read.csv(here::here('data/compiled/tracing_outputs/devphotodraw_tracing_all_tracings.csv')) %>%
  as_tibble() %>%
  mutate(site = str_split_fixed(filename,'_',4)[,1]) %>%
  mutate(session_id = str_split_fixed(filename,'_',4)[,3])
```

```{r}
length(unique(all_data$session_id))
```

```{r}
# predict on new data
ord.pred = predict(ord_wid, all_data, re_formula = NA)
ord.pred = data.frame(ord.pred)
all_data$rating = round(ord.pred$P.Y...1*1 + ord.pred$P.Y...2*2 + ord.pred$P.Y...3*3 + ord.pred$P.Y...4*4)
write.csv(all_data, here::here('data/compiled/tracing_outputs/transformed_tracings.csv')) 
```