### 2021-06-17 analysis/step4_get_features_and_classifications/2_run_classsificatinos/count_classification_outputs.Rmd
for merging individual machine classifications

---
title: "Count and compile classifications"
output: html_notebook
---

## Check we have values for all images
```{r}
library(tidyverse)
library(here)

```

```{r}
current_dir <- here::here("data/compiled/classification-outputs-062021_C_0.1/")

files = dir(current_dir)
indexes = array()
count=0
for (f in files){
  count = count +1
  this_csv = read.csv(paste0(current_dir,'/',f)) 
  indexes[count] = this_csv$index
}
```

```{r}
index_full_array = seq.int(0,max(indexes),1)
index_full_array = as.tibble(index_full_array)

indexes_missing <- index_full_array %>%
  filter(! value %in% indexes)
```

## Make joined CSV
```{r}

files <- dir(current_dir) 
indexes = array()
options(warn=-1) # lots of warningss for full_join over and over even though it's fine

count=0
for (f in files){
  count = count +1
  this_csv = read.csv(paste0(current_dir,'/',f)) 
  if (count==1){
    all_data = this_csv
  }
  all_data <- all_data %>%
    full_join(this_csv)
}
```

```{r}
write.csv(all_data, here::here("data/compiled/", paste0('compiled_classifications/Classification_Outputs',count,'.csv')))

```