### 2021-06-24 analysis/step6_main_analyses/devphotodraw_main_analysis.Rmd
add rmarkdown manuscript file to repo

---
title: "Devphotodraw_main_anlayses"
author: "Bria Long"
date: "6/16/2021"
output:
  html_document:
    toc: true
    theme: united
---
# Preprocessing
## Load libraries
```{r}
library(tidyverse)
library(here)
library(assertthat)
library(langcog)
library(ggthemes)
library(knitr)
library(lmerTest)
library(lme4)

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## Load data
### Compiled classification data
```{r load-classifications}
classification_data <- read.csv(here::here('data/compiled/compiled_classifications/Classification_Outputs2760.csv')) %>%
  as_tibble() %>%
  mutate(age_numeric = age) %>%
  mutate(age = paste('age',age,sep="")) %>%
  mutate(age = as.factor(age)) %>%
  mutate(category = target_label) %>% 
  dplyr::select(-X, -X.1, -index) %>%
  mutate(site = case_when( is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "THU", 
                           !is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "CDM")) %>%
  mutate(site = as.factor(site)) 
```

### Load metadata
```{r}
# load and clean up "category" for tracing trials
all_meta <- read.csv(here::here('data/compiled/metadata/final_merged_metadata.csv')) %>%
    mutate(category = as.character(category)) %>%
    mutate(category = case_when(category == 'this square' ~ 'square',
                                category == 'this shape' ~ 'shape',                                
                                TRUE ~ as.character(category)))

# subIDs are unique identifier in recognition data for THU, session_ids in CDM data
meta_thu <- all_meta %>%
  filter(site=='THU') %>%
  mutate(unique_ids = subID) %>%
  mutate(draw_duration = draw_duration / 1000) # in ms for thu, whoops

meta_cdm <- all_meta %>%
  filter(site=='CDM') %>%
  mutate(unique_ids = session_id)

## unique IDs are now the identifier which will be used to join with machine/human recognition data
all_meta_cleaned <- meta_cdm %>%
  full_join(meta_thu) 
```

### Load tracing data, clean, and avg across participants
```{r}
# import, standardized spliced session_ids, and join back together
# tracing IDs are read in by session_ids, so need these to join with metadata
tracing_thu <- read.csv(here::here('data/compiled/tracing_outputs/transformed_tracings.csv'))  %>%
  mutate(session_id =  paste0('Tsinghua_photodraw_',session_id)) %>%
  filter(site=='Tsinghua') %>%
  left_join(meta_thu, by=c('session_id','category')) %>%
  dplyr::select(-X.1, -X, -site.x, -filename.y) %>%
  rename(site = site.y, filename = filename.x)

tracing_cdm <- read.csv(here::here('data/compiled/tracing_outputs/transformed_tracings.csv'))  %>%
  filter(site=='CDM') %>%
  mutate(session_id = paste0('photodraw_',session_id)) %>%
  left_join(meta_cdm, by=c('session_id','category')) %>%
  dplyr::select(-X.1, -X, -site.x, -filename.y) %>%
  rename(site = site.y, filename = filename.x)

# for modeling tracing scores for each shape (square/shape) separately for each participant
all_tracing <- tracing_thu %>%
  full_join(tracing_cdm)

# for per-subject tracing estimates
tracing_by_sub <- all_tracing %>%
  group_by(unique_ids) %>%
  summarize(avg_tracing_score = mean(rating))

```

### Join meta/tracing with classification data 
```{r}
d <- classification_data %>%
  mutate(unique_ids = session_id) %>%
  left_join(all_meta_cleaned %>% dplyr::select(unique_ids, category, condition, num_strokes, draw_duration, mean_intensity)) %>% # need to select columns so we don't get join errors 
  left_join(tracing_by_sub) %>%
  mutate(age_numeric = as.double(age_numeric)) %>%
  mutate(age_numeric = case_when(age_numeric == 10 ~ 9, # merge 2 10-year-olds into THU data for 9-year-olds
                                 TRUE ~ age_numeric))

```

### Load human classification data
```{r}
humans <- read.csv(here::here('data/compiled/recognition_ratings/compiled_human_recognition.csv'))

humans_merged <- humans %>%
  group_by(unique_ids, category, image_name_short) %>% # group by each drawing of each category
  summarize(num_correct = mean(correct_or_not)) %>%
  left_join(d, by=c('unique_ids','category')) %>% # join with classification data 
  filter(!is.na(age))
```

### Clean up: Get rid of tech error drawings (see note)
```{r}
# bizarrely, there are a few drawings from THU for which stroke data didn't save -- technical error.
# this was checked directly in mongodb database -- no idea why, assuming bad internet.
tech_error_drawings <- d %>%
  filter(is.na(num_strokes))
# filter out tech error drawings from dataset (24)
d <- d %>%
  filter(!is.na(num_strokes)) 
```

```{r}
# this sub did one practice trial, not in subject log or classification data because no category level trials
# missing_id <- meta_cdm %>%
#   filter(!unique_ids %in% d$unique_ids)
```

### Sannity checks on counts of draiwngs per site/condition/etc
```{r}
sanity_counts <- d %>%
  group_by(condition,age_numeric, site) %>%
  summarize(num_drawings = n()) %>%
  kable()
```

```{r}
participant_counts <- d %>%
  group_by(age_numeric, site) %>%
  summarize(num_participants = length(unique(session_id)))

count_by_site <- d %>%
  group_by(site) %>%
  summarize(num_participants = length(unique(session_id)))

count_by_category <- d %>%
  group_by(category) %>%
  summarize(num_drawings = length(unique(session_id)))

count_by_id <- d %>%
  group_by(session_id) %>%
  summarize(num_drawings = length(unique(category))) 
```

265 participants were recruited from the San Jose Children’s Discovery museum, the Palo Alto Junior Museum and Zoo, and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. We aimed to recruit approximately 120 children between 4-9 years of age after exclusions (i.e. 20 4-year-olds, 20 5-year-olds, etc.). In the US-based sample, 135 children participated; 6 participants were excluded, (3) for skipping more than 6 trials, and (3) for scribbling three or more times in a row; (6) participants were tested but their data was not recorded due to a technical error, and (2) participants never made it past the practice trials, leading to a final sample of 121 children. In the China based sample, `r count_by_site$num_participants[2]` children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. A complete breakdown of the number of subjects in each age group and site can be found in the Appendix, Table 1A.

Blank drawings were excluded from analysis, and drawings were randomly undersampled to have an equal number of drawings for each leave-one-out-classification. On average, each child contributed `r mean(count_by_id$num_drawings)` to analysis (min `r min(count_by_id$num_drawings)`, max = `r max(count_by_id$num_drawings)`).

# Descriptive anlayses
## Plot avg human vs. vgg classification
```{r}
ggplot(humans_merged, aes(x=correct_or_not, y=num_correct)) +
  geom_jitter(alpha=.2, width=.1, height=.1) +
  geom_smooth(method='lm') +
  ylab('human recognition') +
  xlab('model recognition')+
  theme_few()
```

```{r}
by_category <- humans_merged %>%
  group_by(category) %>%
  summarize(humans_pc = mean(num_correct), model_pc = mean(correct_or_not, na.rm=TRUE))

ggplot(by_category, aes(x=model_pc, y=humans_pc, label=category)) +
  geom_point() +
  geom_smooth(method='lm') +
  ylab('human recognition') +
  xlab('model recognition') +
  theme_few() +
  ggrepel::geom_label_repel()
```

### Correlates humans vs vgg classifications 
```{r}
cor.test(humans_merged$num_correct, humans_merged$target_label_prob)
cor.test(humans_merged$num_correct, humans_merged$correct_or_not)
```

```{r}

category_age_cond <- humans_merged %>%
  group_by(category, condition) %>%
  summarize(humans_pc = mean(num_correct), model_pc = mean(correct_or_not, na.rm=TRUE))

```

```{r}
ggplot(category_age_cond, aes(x=model_pc, y=humans_pc, color=condition))+
  geom_point() +
  ylim(0,1) +
  xlim(0,1) +
  theme_few() +
  theme(aspect.ratio = 1) +
  facet_wrap(~category) +
  geom_abline(color='grey')

```
```{r}
ggplot(category_age_cond, aes(x=model_pc, y=humans_pc, label=category))+
  geom_point() +
  ylim(0,1) +
  xlim(0,1) +
  theme_few() +
  theme(aspect.ratio = 1) +
  facet_wrap(~condition) +
  geom_abline(xintercept=0, color='grey') +
  ggrepel::geom_label_repel() 
```
## Main descriptive anlayses 
### Compute accuracy by age, condition, site 
```{r descriptives-across-age}
cor_by_age_by_site_by_cond <- d %>%  
  group_by(unique_ids,category,condition,age_numeric, site) %>%
  summarize(avg_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_cor") %>%
  mutate(who = "model")

cor_by_age_by_site_by_cond_humans <- humans_merged %>%  
  group_by(unique_ids,category,condition,age_numeric, site) %>%
  summarize(avg_cor = mean(num_correct)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_cor") %>%
  mutate(who = 'humans')

humans_and_models <- cor_by_age_by_site_by_cond %>%
  full_join(cor_by_age_by_site_by_cond_humans)
```

```{r descriptives-across-age}
cor_by_age_by_cond <- d %>%  
  group_by(unique_ids,category,condition,age_numeric) %>%
  summarize(avg_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_cor") %>%
  mutate(who = "model")

cor_by_age_by_cond_humans <- humans_merged %>%  
  group_by(unique_ids,category,condition,age_numeric) %>%
  summarize(avg_cor = mean(num_correct)) %>%
  group_by(age_numeric,condition) %>%
  multi_boot_standard(col = "avg_cor") %>%
  mutate(who = 'humans')

humans_and_models_no_site <- cor_by_age_by_cond_humans %>%
  full_join(cor_by_age_by_cond)
```

### Plot accuracy by age, condition, site 
```{r}
base_size_chosen=12; smooth_alpha=.2
ggplot(humans_and_models, aes(age_numeric,mean, col = condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) +
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  ylim(0,1) + 
  geom_smooth(span=10, alpha=.2) +
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_grid(who~site)
```
```{r}
base_size_chosen=12; smooth_alpha=.2
ggplot(humans_and_models_no_site, aes(age_numeric,mean, col = conditionf)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) +
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  ylim(0,1) + 
  geom_smooth(span=10, alpha=.2) +
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_grid(~who)

```
```{r}
disagreements <- humans_merged %>%
  filter(num_correct>.8) %>%
  filter(correct_or_not==0) %>%
  filter(condition=='S')

disagreements <- disagreements %>%
  mutate(image_path = here::here('data/compiled/drawings/object_drawings', image_name_short)) %>%
  mutate(new_image_path = here::here('data/compiled/drawings/disagreements', image_name_short))

dir.create(here::here('data/compiled/drawings/disagreements'))

file.copy(disagreements$image_path, disagreements$new_image_path)

```

```{r}
machine_p_advantage <- humans_merged %>%
  filter(num_correct<.5) %>%
  filter(correct_or_not==1) %>%
  filter(condition=='P')

machine_p_advantage <- machine_p_advantage %>%
  mutate(image_path = here::here('data/compiled/drawings/object_drawings', image_name_short)) %>%
  mutate(new_image_path = here::here('data/compiled/drawings/machine_p_advantage', image_name_short))

dir.create(here::here('data/compiled/drawings/machine_p_advantage'))

file.copy(machine_p_advantage$image_path, machine_p_advantage$new_image_path)

```

### Compute effort/tracing by age, condition, site 
```{r}
draw_duration <- d %>%
  group_by(unique_ids,condition,age_numeric, site) %>%
  summarize(avg_draw_duration = mean(draw_duration)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_draw_duration")

num_strokes <- d %>%
  group_by(unique_ids,condition,age_numeric,site) %>%
  summarize(avg_num_strokes = mean(num_strokes)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_num_strokes")

avg_intensity <- d %>%
  group_by(unique_ids,condition,age_numeric,site) %>%
  summarize(avg_intensity = mean(mean_intensity)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_intensity")

tracing_scores <- d %>%
  distinct(avg_tracing_score, unique_ids, site, age_numeric) %>%
  group_by(age_numeric, site) %>%
  multi_boot_standard(col = "avg_tracing_score")

```

### Plot effort/tracing by age, condition, site 
```{r plot-descriptives-across-age}
## Make compiled plot of descriptives
base_size_chosen=18 # size of text in plots
smooth_alpha=.2

p1=ggplot(draw_duration, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Draw duration (s)') +
  theme(legend.position = "right") + 
  ylim(0,30) +
  facet_grid(~site)

p2=ggplot(avg_intensity, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Ink used (mean intensity)') +
  theme(legend.position = "right") + 
  ylim(.02,.08) +
  facet_grid(~site)

p3=ggplot(num_strokes, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Number of strokes') +
  theme(legend.position = "right") +
  ylim(0,30) +
  facet_grid(~site)
```

### Plot effort covariates
Looks like THU kids spent more time on drawings, interesting; otherwise no big differences.

```{r}
p1
```

```{r}
p2
```

```{r}
p3
```

### Plot traing scores
```{r}
ggplot(tracing_scores, aes(age_numeric,mean)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Tracing score') +
  theme(legend.position = "none") +
  geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
  facet_grid(~site) +
  ylim(0,4)
```

## Correlation of item effects across sites
Wow, cats are always badly classified.
```{r}
item_effects <- d %>%
  group_by(session_id,age_numeric,category, site) %>%
  summarize(avg_sub_cor = mean(correct_or_not)) %>%
  group_by(age_numeric,category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects, aes(x=CDM, y=THU, color=age_numeric)) +
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  ylim(0,1) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_wrap(~category)
```

```{r}
item_effects <- humans_merged %>%
  group_by(session_id,age_numeric,category, site) %>%
  summarize(avg_sub_cor = mean(num_correct)) %>%
  group_by(age_numeric,category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects, aes(x=CDM, y=THU, color=age_numeric)) +
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  ylim(0,1) + 
  xlim(0,1) +
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  geom_abline(color = 'grey', intercept=0) + 
  facet_wrap(~category)
```

```{r}
item_effects_not_spread <- humans_merged %>%
  group_by(session_id,age_numeric,category, site) %>%
  summarize(avg_sub_cor = mean(num_correct)) %>%
  group_by(age_numeric,category, site) %>%
  multi_boot_standard(col = "avg_sub_cor")

ggplot(item_effects_not_spread, aes(x=as.numeric(age_numeric), y=mean, color=site)) +
  geom_point() +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) +
  theme_few(base_size = base_size_chosen) + 
  ylim(0,1) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  ylab('Proportion correct') +
  xlab('Age (in years)') +
  geom_abline(color = 'grey', intercept=0) +
  geom_smooth(span=20, alpha=.1) +
  facet_wrap(~category)
```

```{r}
item_effects <- humans_merged %>%
  group_by(age_numeric,category) %>%
  summarize(model_pc = mean(correct_or_not), human_pc = mean(num_correct))

ggplot(item_effects, aes(x=model_pc, y=human_pc, color=age_numeric)) +
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  ylim(0,1) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_wrap(~category)
```

```{r}
item_effects_no_age <- d %>%
  group_by(session_id,category, site) %>%
  summarize(avg_sub_cor = mean(correct_or_not)) %>%
  group_by(category, site) %>%
  summarize(avg_cor = mean(avg_sub_cor)) %>%
  spread(key=site, value=avg_cor)

ggplot(item_effects_no_age, aes(x=CDM, y=THU, color=category)) +
  geom_point() + 
  theme_few(base_size = base_size_chosen) + 
  geom_abline(intercept = 0, slope=1)
```

# Inferential stats
## Try maximumal model first
Singular boundary fits when including (condition | subject) random slopes
Won't converge with random slopes for condition x age xs ite on each category
```{r}
full_model <- glmer(correct_or_not ~ condition*scale(age_numeric)*site +
                        (1 | unique_ids) +
                        (condition*scale(age_numeric) | category),
      data = d, family="binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

summary(full_model)
```

```{r}
full_model_humans <- lmer(num_correct ~ condition*scale(age_numeric)*site +
                        (1 | unique_ids) +
                        (condition | category),
      data = humans_merged)
summary(full_model_humans)
```

### Remove 3-way interaction
No interactions with site are significant, removing from main effects structure
When 3 way interaction is removed, still can't add back more random slopes.
```{r}

full_model_no_int <- glmer(correct_or_not ~ condition*scale(age_numeric)+site +
                        (1 | unique_ids) +
                        (condition*scale(age_numeric) | category),
      data = d, family="binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

summary(full_model_no_int)

```

```{r}
full_model_no_int_humans <- lmer(num_correct ~ condition*scale(age_numeric)+site +
                        (1 | unique_ids) +
                        (condition | category),
      data = humans_merged)

summary(full_model_no_int_humans)

```

ANOVA suggests that these two models -- with/without 3-way int -- explain equal variance.
```{r}
anova(full_model, full_model_no_int)
```

### Add back site*condition
```{r}
with_site_condition <- glmer(correct_or_not ~ condition*scale(age_numeric)+condition*site +
                        (1 | unique_ids) +
                        (condition*scale(age_numeric) | category),
      data = d, family="binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

summary(with_site_condition)

```

### Add back age*site
```{r}
with_age_site <- glmer(correct_or_not ~ condition*scale(age_numeric)+scale(age_numeric)*site +
                        (1 | unique_ids) +
                        (condition*scale(age_numeric) | category),
      data = d, family="binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

summary(with_age_site)

```

## Add back both
```{r}

with_site_condition_and_site_age <- glmer(correct_or_not ~ condition*scale(age_numeric)+condition*site+scale(age_numeric)*site+
                        (1 | unique_ids) +
                        (condition*scale(age_numeric) | category),
      data = d, family="binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

summary(with_site_condition_and_site_age)
```

## Add in effort/tracing covariates — does effort mediate any of the differences we're seeing? Not really. Still see same main effects structure.
```{r}
full_model_with_effort <- glmer(correct_or_not ~ condition*scale(age_numeric)*site +
                      scale(avg_tracing_score) +
                      scale(mean_intensity) +
                      scale(draw_duration) +
                      scale(num_strokes) + 
                      (1 | unique_ids) +
                      (condition*scale(age_numeric)  | category),
      data = d, family="binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# frustratingly can't get anything with random slopes to converge!
summary(full_model_with_effort)

```

## Repeat with human classificatinos
```{r}
full_model_with_effort <- lmer(num_correct ~ condition*scale(age_numeric)*site +
                      scale(avg_tracing_score) +
                      scale(mean_intensity) +
                      scale(draw_duration) +
                      scale(num_strokes) + 
                      (1 | unique_ids) +
                      (condition |category), data = humans_merged)

# condition*age slopes on categories won't converge
summary(full_model_with_effort)
```

## No difference in tracing abilities across sites; only across age
```{r}
tracing_diff = lmer(rating ~ age_numeric + site + (1|unique_ids), data = all_tracing)
summary(tracing_diff)
```