### 2021-11-13 writing/Manuscript_June2021.Rmd
up to end of methods!

---
title             : "Developmental consistency in children’s drawings of object categories"
shorttitle        : "Development of drawing"

author: 
  - name          : "Bria Long"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Jane Stanford Way, Stanford CA 94305"
    email         : "bria@stanford.edu"
  - name          : "Ying Wang"
    affiliation   : "2"
  - name          : "Stella Christie"
    affiliation   : "2"
  - name          : "Michael C. Frank"
    affiliation   : "1"
  - name          : "Judith E. Fan"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "Stanford University"
  - id            : "2"
    institution   : "Tsinghua University"
  - id            : "3"
    institution   : "University of California, San Diego"

authornote: |

abstract: |
 Childen's drawings of common object categories become dramatically more recognziable across development. What are the major factors that explain this developmental change? Here we examined the degree to which these developmental changes in recognizibility vary across different drawing tasks (i.e. drawing from observation vs. from memory), geographical locations (San Jose, US vs. Beijing, China), and with children's tracing abilities. To do so, we collecting digital drawings of object categories (e.g., cat, airplane) from 4-9 year-olds (N=253). Overall, we show broad consistency in these developmental trajectories across both drawing tasks and these two geographical locations, and we find that children's tracings abilites are good predictors of the recognizability of the drawings that they produce. In addition, we find that our Beijing sample produced more recognizable drawings but showed similar tracing abilities to childen from San Jose, USA. Overall, this work suggests that the developmental trajectory of children's drawings are remarkably consistent and not easily explainable by changes in visuomotor control or working memory capacity.

keywords          : "children's drawings, visual production, tracing, object recognition, visuomotor control"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(tidyverse)
library(here)
library(assertthat)
library(langcog)
library(ggthemes)
library(knitr)
library(lmerTest)
library(lme4)
library(xtable)
library(viridis)
```

```{r}

knitr::opts_chunk$set(fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
```

As humans, we have many powerful tools to externalize what we perceive and know, including language and gesture.
One tool that has been especially transformative for human cognition and culture is graphical representation, which allows people to encode their thoughts in a visible, durable format.
Drawing is an important case study in graphical representation, being a technique that dates back 60,000 years [@hoffmann2018u], well before the emergence of symbolic writing systems, and is practiced in many cultures.

In modern times, drawings are produced prolifically by children from an early age. 
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities [@minsky1972artificial], and accordingly a long history of work has examined changes in how children draw objects across development [@piaget1929child; @kellogg1969analyzing; @karmiloff1990constraints; @fury1997children].
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age: younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between target concepts (e.g., \textit{adult} vs. \textit{child}) than older children, who enrich their drawings with more diagnostic part [@sitton1992drawing] and relational [@light1983effects] information.

What drives these dramatic changes in children's drawings across development? 
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements [@freeman1987current; @rehrig2018does]. 
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints on children's drawings — including what children know about a given target object and how well children are able to access this conceptual knowledge when producing a drawing.

Prior work has provided evidence that changes in children's drawings partly reflect changes in children's mental representations of visual concepts. In a large observational dataset, older children produced drawings of object categories that were more diagnostic of the categories they were trying to depict [@long2018drawings; @long2021parallel]. This result held even when accounting for differences in basic shape tracing abilities and the amount of effort children expended on individual drawings. Furthermore, older children also relied more on these same diagnostic visual features when recognizing other children's drawings.
Together, these results suggest that developmental change in these underlying conceptual representations drive parallel changes in both children's production and recognition of drawings, and thus that change in how well children can produce recognizable drawings do not only reflect improvements in motor control [@natu2016development; @dekker2011dorsal].

However, this work leaves open the contribution of children's evolving abilities to retrieve and maintain relevant information in memory. Children's ability to access semantic knowledge about categories and to maintain this information in mind when producing a drawing likely changes across childhood [@pailian2016visual]. Here, we directly test the idea that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant visual features of different categories: that is, when asked to "draw a [rabbit]", they may struggle to conjure up the relevant visual details and then hold in mind what rabbits tend to look like. On this account, providing children with additional visual information about different categories -- for example, via canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults [@fan2018common].

However, prior work also suggests that younger children tend to draw what they know about objects rather than integrate information in their immediate perceptual experience [@luquet1927dessin]. For example, when asked to draw from observation, younger children tend to include features that are not visible from their vantage point, yet are diagnostic of category membership (e.g., a handle on a \textit{mug}) [@bremmer1984prior; @barrett1976symbolism], and only omit these features later in development. Similarly, young children will insist that their nearly identical drawings of different concepts (e.g., balloon and person) unambiguously refer to different things [@bloom1998intention]. 
Thus, an alternative possibility is that only older children may be able to produce more recognizable drawings when provided with canonical exemplars of different categories. 
On this account, memory constraints are not a major factor that drives developmental changes in children's drawings.

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when children were provided with a verbal cue ("Can you draw a [rabbit]?) versus when provided with a picture cue ("Can you draw this [rabbit] as it looks in the picture?"). On verbal-cue trials, children thus must access their mental representation of a "rabbit" and choose the features necessary to convey that object's identity. Conversely, on picture-cue trials, children are explicitly asked to rely on the visual features provided in a canonical photograph of each object category.

To test the generality of our findings, we recruited children from two sites in different countries — San Jose, USA and Beijing, China.
Most empirical studies on children's drawings have been conducted exclusively on small samples of children from the United States or Western Europe, limiting their generalizability. Further, children in different communities may spend more or less time practicing drawing or use different visual conventions to draw [@cohn2012explaining; @willats2006making]. While some prior work has focused on differences in children's drawings across different countries, most of this work has focused on differences in educational practices that are thought to manifest in children's drawings [@la2001children; @winner1989can; @huntsinger2011cultural]. Leaving aside questions about potential causes of differences between sites, here we focus on measuring the degree to which drawing abilities vary between children at these two sites using the same experimental protocol.

Finally, we assessed the degree to which visuomotor development accounts for observed developmental changes in drawing recognizability. While it is uncontroversial that visuomotor control can constrain how and what we can draw, little work has directly related measures of visuomotor control to measures of drawing recognizability [@long2021parallel]. To do so, we measured each child's visuomotor control via a shape tracing task and related these measurements of tracing accuracy to the recognizability of the drawings that each child produced.

In sum, in the current study we collect shape tracings and digital drawings of visual concepts from 4-9 year-old children in Beijing, China and San Jose, USA using both picture cues and verbal cues. In doing so, we make three key contributions to our understanding of the development of children's drawings. First, we replicate our prior findings  [@long2021parallel] that the recognizability of children's drawings increases steadily throughout this age range (4-9 years). Second, we test the degree to which memory constraints might account for these developmental changes. In keeping with accounts of naïve realism and against a large role of limitations on memory, we predicted that only older children would be able to use the visual information present in the canonical photographs to improve their drawings. Third, we test the generalizability of our findings across two sites in different countries. We predicted that we would see convergence in the development of drawing abilities across both geographical sites, with older children becoming progressively better at producing recognizable drawings. In addition, we predicted that most of the variance across geographical sites in drawing ability would be explainable by differences in visuomotor control, operationalized as performance on a shape tracing task; these primary analyses were pre-registered at: https://osf.io/qymjr/.

# Methods
```{r plotting-params}
base_size_chosen=12; smooth_alpha=.2
```

```{r load-classifications}
classification_data <- read.csv(here::here('data/compiled/compiled_classifications/Classification_Outputs2760.csv')) %>%
  as_tibble() %>%
  mutate(age_numeric = age) %>%
  mutate(age = paste('age',age,sep="")) %>%
  mutate(age = as.factor(age)) %>%
  mutate(category = target_label) %>% 
  dplyr::select(-X, -X.1, -index) %>%
  mutate(site = case_when( is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "THU", 
                           !is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "CDM")) %>%
  mutate(site = as.factor(site)) 
```

```{r load-metadata}
# load and clean up "category" for tracing trials
all_meta <- read.csv(here::here('data/compiled/metadata/final_merged_metadata.csv')) %>%
    mutate(category = as.character(category)) %>%
    mutate(category = case_when(category == 'this square' ~ 'square',
                                category == 'this shape' ~ 'shape',                                
                                TRUE ~ as.character(category)))

# subIDs are unique identifier in recognition data for THU, session_ids in CDM data
meta_thu <- all_meta %>%
  filter(site=='THU') %>%
  mutate(unique_ids = subID) %>%
  mutate(draw_duration = draw_duration / 1000) # in ms for thu, whoops

meta_cdm <- all_meta %>%
  filter(site=='CDM') %>%
  mutate(unique_ids = session_id)

## unique IDs are now the identifier which will be used to join with machine/human recognition data
all_meta_cleaned <- meta_cdm %>%
  full_join(meta_thu) 
```

```{r load-tracing}
# import, standardized spliced session_ids, and join back together
# tracing IDs are read in by session_ids, so need these to join with metadata
tracing_thu <- read.csv(here::here('data/compiled/tracing_outputs/transformed_tracings.csv'))  %>%
  mutate(session_id =  paste0('Tsinghua_photodraw_',session_id)) %>%
  filter(site=='Tsinghua') %>%
  left_join(meta_thu, by=c('session_id','category')) %>%
  dplyr::select(-X.1, -X, -site.x, -filename.y) %>%
  rename(site = site.y, filename = filename.x)

tracing_cdm <- read.csv(here::here('data/compiled/tracing_outputs/transformed_tracings.csv'))  %>%
  filter(site=='CDM') %>%
  mutate(session_id = paste0('photodraw_',session_id)) %>%
  left_join(meta_cdm, by=c('session_id','category')) %>%
  dplyr::select(-X.1, -X, -site.x, -filename.y) %>%
  rename(site = site.y, filename = filename.x)

# for modeling tracing scores for each shape (square/shape) separately for each participant
all_tracing <- tracing_thu %>%
  full_join(tracing_cdm)

# for per-subject tracing estimates
tracing_by_sub <- all_tracing %>%
  group_by(unique_ids) %>%
  summarize(avg_tracing_score = mean(rating))
```

```{r join-data}
model_classifications <- classification_data %>%
  mutate(unique_ids = session_id) %>%
  left_join(all_meta_cleaned %>% dplyr::select(unique_ids, category, condition, num_strokes, draw_duration, mean_intensity)) %>% # need to select columns so we don't get join errors 
  left_join(tracing_by_sub)

# bizarrely, there are a few drawings from THU for which stroke data didn't save -- technical error.
# this was checked directly in mongodb database -- no idea why, assuming bad internet.
tech_error_drawings <- model_classifications %>%
  filter(is.na(num_strokes))
# filter out tech error drawings from dataset (24)
model_classifications <- model_classifications%>%
  filter(!is.na(num_strokes))

```

```{r load-human-recognition}
humans <- read.csv(here::here('data/compiled/recognition_ratings/compiled_human_recognition.csv'))

```

```{r join-human-rec-with-meta}
d <- humans %>%
  group_by(unique_ids, category, image_name_short, site, age) %>% # group by each drawing of each category
  summarize(num_correct = sum(correct_or_not==TRUE), num_incorrect = sum(correct_or_not==FALSE), prop_correct = mean(correct_or_not))  %>%
   left_join(all_meta_cleaned %>% dplyr::select(unique_ids, category, condition, num_strokes, draw_duration, mean_intensity)) %>%
  mutate(age_numeric = age) %>%
  mutate(age_numeric = as.double(age_numeric)) %>%
  mutate(age_numeric = case_when(age_numeric == 10 ~ 9, # merge 2 10-year-olds into THU data for 9-year-olds
                                 TRUE ~ age_numeric)) %>%
  left_join(tracing_by_sub) %>%
  mutate(site = fct_recode(site, "San Jose" = "CDM", "Beijing" = "THU"))%>%
  mutate(condition = fct_recode(condition, "Picture cue" = "P", "Verbal cue" = "S"))

humans_and_models_merged <- humans %>%
  group_by(unique_ids, category, image_name_short) %>% # group by each drawing of each category
  summarize(prop_correct = mean(correct_or_not)) %>%
  left_join(model_classifications, by=c('unique_ids','category')) %>% # join with classification data 
  filter(!is.na(age))
```

```{r count-participants}
participant_counts <- d %>%
  group_by(age_numeric, site) %>%
  summarize(num_participants = length(unique(unique_ids)))

count_by_site <- d %>%
  group_by(site) %>%
  summarize(num_participants = length(unique(unique_ids)))

count_by_category <- d %>%
  group_by(category) %>%
  summarize(num_drawings = length(unique(unique_ids)))

count_by_id <- d %>%
  group_by(unique_ids) %>%
  summarize(num_drawings = length(unique(category))) 
```

## Participants
265 participants were recruited from two local children's museum in [BLINDED] and at preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e. 20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of `r count_by_site$num_participants[1]` children. In the Beijing sample, `r count_by_site$num_participants[2]` children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed `r mean(count_by_id$num_drawings)` drawings to analysis (min `r min(count_by_id$num_drawings)`, max = `r max(count_by_id$num_drawings)`). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at [blinded] (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at [blinded] in Beijing, China.

## Stimuli and Task Procedure

Children were seated in front of a touchscreen tablet (Ipad Pro 12.9") during each drawing session.
At the beginning of the session, a trained experimenter first told each child, “After this game is over, someone is going to try to recognize what you were trying to draw. So, please draw so that someone else could guess what you were trying to draw.” 
A native English speaker gave these instructions to the San Jose sample, and a native Mandarin speaker gave a translation of these instructions to the Beijing sample. 
Following these instructions, children completed 2 shape-tracing trials, providing a baseline measure of visuomotor control in the absence of memory demands.
All children first traced a square, then a more complex shape (see Figure \ref{fig:example-tasks}).
After these tracing trials, children produced drawings of 12 familiar object categories (airplane, bike, bird, car, cat, chair, cup, hat, house, rabbit, tree, watch).
Strokes could not be deleted once drawn, and children had to complete each trial within 30 seconds.
Across trials, we manipulated the type of cue (verbal vs. picture) children received before producing each drawing (Fig. \ref{fig:example-tasks}), providing a measure of the impact of reminding children of what typical exemplars of each category looked like. 
Specifically, children first completed six trials with one cue type, then switched to the other cue type. 
The order in which each cue type was used was counterbalanced across children.
On verbal-cue trials, children viewed a short video clip in which an experimenter named the target category: “What about a [cat]? Can you draw a [cat]?”.  
On picture-cue trials, children viewed a photograph of a typical exemplar of the target category while listening to an audio clip of the same experimenter who said, “What about this [cat]? Can you draw the [cat] as it looks in the picture?”
The photograph remained on the screen for the duration of the drawing trial.
There were three different exemplars of each category used in this study, one of which was randomly sampled on each picture-cue trial.
All experimental code, videos, translations, and stimuli are available on the public repository for this project.

```{r example-tasks, fig.env = "figure", out.width="100%", fig.pos = "H", fig.cap = "Example trials from the tracing assessment and two drawing tasks."}
knitr::include_graphics("figs/task_examples.png")
```

## Measuring drawing effort covariates
We recorded both the final drawings and the location of each stroke in a digital format, allowing us to precisely measure several different variables that provided proxies for the amount of effort children invested in producing each drawing.
Specifically, we measured: (1) the amount of time spent (i.e., end time of last stroke — start time of first stroke), (2) the number of individual strokes drawn, and (3) the amount of "ink" used (i.e., proportion of drawing canvas that was filled).

## Measuring Tracing Accuracy
As in prior work [@long2021parallel], we used a semi-automated procedure for evaluating how accurately each child performed the tracing task that was validated against empirical judgments of tracing quality. In brief, we decomposed tracing accuracy into two components: a shape error component and a spatial error component. Shape error reflects how closely the participant’s tracing matched the contours of the target shape; the spatial error reflects how closely the location, size, and orientation of the participant’s tracing matched the target shape. These two error components were computed automatically for each tracing and used to yield a "tracing score" for each tracing that mirrors adult human judgments of tracing quality (see Appendix for details).

## Measuring drawing recognizability

### Human recognition scores
We measured the recognizability of each drawing via an online recognition experiment. 
Adult participants based in the U.S. were recruited via Prolific for a 15-minute experiment, compensated at $14/hour, and asked to identify the category depicted in a random subset of approximately 140 drawings. 
Each subset of drawings was balanced with respect to age, category, and site, and each drawing was shown to 10 participants. 
Participants were shown these drawings in a random sequence and asked "What does this look like?"
Participants selected their responses from the set of 12 categories and were encouraged to provide their best guess if they were unsure. 
No participants were excluded from analysis for missing the catch trial, which was included to verify that participants could accurately describe their goal in this task. 
We then computed a recognition score for each drawing, reflecting the proportion of participants who correctly identified the target category.

### Automated recognition scores
We used a combination of deep convolutional neural network (CNN) features and logistic regression to obtain automated recognition scores, per our preregistered protocol. 
To encode the high-level visual features of each drawing, we passed each drawing into VGG-19 [@simonyan2014very], a deep CNN architecture that had been pre-trained on Imagenet classification and implemented in PyTorch [@paszke2017automatic].
We extracted the feature representation of this drawing in the second-to-last layer of this network (FC6). 
Feature representations in this layer consist of flat 4096-dimensional vectors, to which we applied channel-wise normalization. 
Next, we used these features to train a 12-way logistic classifier with L2 regularization (tolerance = .1, regularization = .1), and used this classifier to predict the category label for each drawing.
To ensure a balanced distribution of drawings appearing in the training set, we randomly undersampled such that there was an equal number of drawings for each combination of geographical site (San Jose, Beijing) and the 12 categories. 
\textbf{jefan query: Here we describe the training data distribution, but not the test set? What kind of train/test splitting procedure did we use here (LOO? 80/20 CV?)}
No additional metadata about the age of the child who produced each sketch was provided to the model. 
This procedure was repeated for each drawing in the dataset, yielding a binary recognition score for each drawing.

## Statistical models
```{r}
human_v_model = cor.test(humans_and_models_merged$prop_correct, humans_and_models_merged$correct_or_not)
```

To evaluate our main hypotheses, we fit generalized linear mixed-effects models to the human recognition scores to assess the factors that influenced the recognizability of the drawings that children produced. A first generalized linear mixed effect model was fit to the recognizability scores for each drawing, including fixed effects of children's age (in years), geographical site (San Jose vs. Beijing), and drawing task (verbal cue vs. picture cue) and the three-way interaction between these key variables. We initially planned to include random slopes for the effect of drawing task on each child (as this varied within-subjects), and random slopes for the effect the full three-way interactions between task, age, and site on each category. However, models with this random-effects structure failed to converge, and the reported models use the maximal random effects structure that did converge, which included random slopes for the two-way interaction between task and age on each category.

In a secondary analysis, we investigated the degree to which any of the above effects were mediated by children's tracing abilities or the amount of effort that children expended while drawings. 
We thus fit the same model above, with the addition of fixed-effect terms for each child's estimated tracing score (see Measuring Tracing Abilities), the time that child spent drawing (in seconds), the amount of "ink" used (i.e. percentage of non-white pixels in the drawn image), and the number of strokes. 
These additional predictors were first z-scored before being included in the model.
Finally, we also assessed the degree to which tracing ability development differed across geographical sites, where tracing scores were modeled as a function of age (in years), site, and their interaction, with the same random-effects structure as the first model.

This analysis plan was pre-registered at https://osf.io/qymjr. While we had initially planned to use automated recognition scores as our main dependent variable, following prior work [@long2021parallel], we found that these automated recognition scores were only modestly correlated with human recognition scores ($r$ = `r round(human_v_model$estimate,2)`, $t$ = `r round(human_v_model$statistic,2)`, p < .001); furthermore, descriptive plots of the automated recognition scores revealed an overall difference between the two drawing tasks that was not evident in the human recognition data. To be conservative, we thus fit all of our models to the human recognition data but have made all model predictions available to facilitate future research.

```{r}
full_model <- glmer(cbind(num_correct, num_incorrect) ~ condition*scale(age_numeric)*site +
                        (condition | unique_ids) +
                        (condition*scale(age_numeric) | category),
      data = d, family='binomial', control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

full_model_out = summary(full_model)

```

```{r}
full_model_with_effort <- glmer(cbind(num_correct, num_incorrect) ~ condition*scale(age_numeric)*site +
                      scale(avg_tracing_score) +
                      scale(mean_intensity) +
                      scale(draw_duration) +
                      scale(num_strokes) + 
                      (condition | unique_ids) +
                      (condition*scale(age_numeric) | category),
      data = d, family='binomial', control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# frustratingly can't get anything with random slopes to converge!
full_model_with_effort_out = summary(full_model_with_effort)

```

```{r}
site_difference <- glmer(cbind(num_correct, num_incorrect) ~ poly(scale(age_numeric),2)*site +
                      (1 | unique_ids) +
                      (1 | category),
      data = d, family='binomial', control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

site_difference_out = summary(site_difference)

```

# Results
### Confirmatory Analyses
Overall, we found relative consistency in the developmental trajectory of children's drawings: Figure \ref{fig:main-results} shows the recognizability of children's drawings at each age as a function of the drawing task and the geographical site that they were located in (all model coefficients can be found in Table 1). We found steady changes in the recognizability of children's drawings across this age range, replicating prior work with an observational dataset collected at a free-standing kiosk [@long2021parallel]. Further, we found that these changes in recognizability did not depend on the drawing task children completed: Children's drawings were equally recognizable when they were asked to "Draw this rabbit as it looks in the picture" versus they were simply asked "Can you draw a rabbit?". While we anticipated that older children might produce more recognizable drawings when provided with a picture cue -- as do adults when presented with canonical photographs of object categories [@yang2021visual] -- this was not the case. However, we did observe a main effect of geographical site: we found that children from Beijing, China overall tended to produce more recognizable drawings than children from San Jose, USA (see coefficients in Table 1). A closer look comparing the recognizability of the drawings made at children at each age and site hints that the children from Beijing, China may tend to produce more recognizable drawings earlier in development.

In a second set of analyses, we aimed to understand the nature of this difference in recognizability across sites. We initially hypothesized that any differences in the recognizability of children's drawings across sites would be mediated by the amount of effort children expended while drawing or their tracing abilities.  We thus first examined how effort and tracing abilities varied across task, age, and geographical location. Figure \ref{fig:effort-covariates} shows three effort covariates -- average intensity, number of strokes used, and time spent drawing -- measured for each drawing as a function of children's age, drawing task, and geographical site.  Children tended to spend more time drawing provided with a picture cue and asked to draw "the [rabbit] as it looks in the picture", and children from Beijing tended to spend more time drawing overall. In contrast, children's tracing abilities did not vary across geographical site: Figure \ref{fig:tracing} shows the average tracing score at each age at each site. While children's tracing abilities increased steadily with age, as in prior work [@long2021parallel], children from both sites produced equally good tracings.

We then included children's average tracing scores and effort covariates measured for each drawing (average intensity, number of strokes used, and time spent drawing) as fixed effects into a second generalized linear mixed-effects model (see Statistical Models). If the amount of effort spent or children's tracing abilities accounted the differences between sites that we observed, we reasoned that we should no longer observe a main effect of geographical location on drawing recognizability. Contra this hypothesis, however, we still observed a significant effect of geographical site (see Table 2), despite the fact that individual children's tracing abilities were strong predictors of how recognizable their drawings were. In addition, the amount of time a child spent drawing was in fact a significant negative predictor of recognizability -- indicating that the amount of effort that children expended did not necessarily result in a drawing that was more recognizable to others. Thus, these results suggest that these two groups of children differ in their ability to produce recognizable drawings of object categories in a way that is not easily explainable by effort covariates or tracing abilities.

### Exploratory Analyses
In a third set of exploratory analyses, we examined how the developmental trajectory of the recognizability of children's drawings differ for different object categories. Intuitively, some object categories (e.g., cat) may be easier to draw than others (e.g., watch), resulting in shallower or steeper changes in recognizability over age. However, we had no strong hypotheses about how these trajectories might additionally vary with geographical locations.

Figure \ref{fig:item-effects} shows the recognizability of children's drawings for each of the 12 categories a function of geographical location and children's age and reveal considerable variability.  For example, certain categories (e.g., cats and rabbits) were overall more difficult for children to depict such that they were distinct from one another, particularly for younger children. In addition, children from Beijing produced more recognizable drawings of certain categories at all ages -- including airplanes, birds, and rabbits -- while older children from San Jose produced more recognizable drawings of bikes.  Figure \ref{fig:example-drawings} shows recognizable, randomly sampled example drawings from 6-year-olds at each category, site, and condition. These exploratory analyses thus reveal systematic differences between object categories and geographical site that do not seem easily attributable to differences in effort.

```{r render-drawings, include=FALSE}
# set.seed(123)
# set.seed(234)
# sample_drawings <- d %>%
#   filter(prop_correct>.8) %>%
#   filter(age_numeric==6) %>%
#   group_by(category, condition, site) %>%
#   sample_n(1)
#   
# sample_drawings <- sample_drawings %>%
#   mutate(image_path = here::here('data/compiled/drawings/object_drawings', image_name_short)) %>%
#   mutate(new_image_path = here::here('data/compiled/drawings/sample_drawings2', image_name_short)) 
# 
# dir.create(here::here('data/compiled/drawings/sample_drawings2'))
# 
# file.copy(sample_drawings$image_path, sample_drawings$new_image_path)

```

```{r check-drawing-time}
# drawing_time <- lmer(draw_duration ~ condition*scale(age_numeric)*site +
#                       (1 | unique_ids) +
#                       (1 | category),
#       data = d)
# drawing_time_out = summary(drawing_time)
```

```{r compile-main-results}
cor_by_age_by_site_by_cond <- d %>%  
  group_by(unique_ids,category,condition,age_numeric, site) %>%
  summarize(avg_cor = mean(prop_correct)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_cor")

cor_by_age_by_site_by_cond_indiv <- d %>%  
  group_by(unique_ids,category,condition,age_numeric, site) %>%
  summarize(avg_cor = mean(prop_correct))
```

```{r main-results, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=5, fig.cap = "Proportion of drawings recognized as a function of the age of the child who completed each drawing, the geographical site they were tested at (San Jose vs. Beijing), and the type of drawing task they completed. Individual data points represent drawings within each condition by an individual participant and are slightly jittered. Error bars show bootstrapped 95 percent cofnidence intervals."}
ggplot(cor_by_age_by_site_by_cond, aes(age_numeric,mean, col = site)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) +
  geom_point(data = cor_by_age_by_site_by_cond_indiv, aes(x=age_numeric, y=avg_cor), position=position_jitterdodge(jitter.width = NULL,
  jitter.height = .001,
  dodge.width = 0.2,
  seed = 123), alpha=.05) +
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age (in years)', y='Proportion drawings recognized') +
  ylim(-.01,1.1) + 
  geom_smooth(span=20, alpha=.1) +
  scale_color_viridis(discrete=TRUE, name="", option='B', begin=0, end=.4) +
  theme(legend.position = 'bottom') +
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_grid(~condition)
```

```{r output-models-1}
cleaned_names_full_model = c("Intercept","Task","Age","Site","Task*Age","Task*Site","Age*Site","Task*Age*Site")
rownames(full_model_out$coefficients)<-cleaned_names_full_model

# xtable(full_model_out$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed mode predicting the recognizability of each drawing for the main experimental contrasts.")
```

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & 1.39 & 0.16 & 8.51 & 0.000 \\ 
  Task & 0.02 & 0.26 & 0.08 & 0.936 \\ 
  Age & 1.16 & 0.14 & 8.59 & 0.000 \\ 
  Site & 0.60 & 0.16 & 3.71 & 0.000 \\ 
  Task*Age & -0.14 & 0.19 & -0.73 & 0.467 \\ 
  Task*Site & -0.19 & 0.16 & -1.17 & 0.244 \\ 
  Age*Site & 0.15 & 0.16 & 0.91 & 0.361 \\ 
  Task*Age*Site & 0.06 & 0.16 & 0.38 & 0.706 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed mode predicting the recognizability of each drawing for the main experimental contrasts.} 
\end{table}

```{r output-models-2}
cleaned_names_full_model_with_effort = c("Intercept","Task","Age","Site","Est. tracing score","Avg. intensity", "Draw duration", "Number of strokes","Task*Age","Task*Site", "Age*Site","Task*Age*Site")

rownames(full_model_with_effort_out$coefficients)<-cleaned_names_full_model_with_effort

# xtable(full_model_with_effort_out$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed mode predicting the recognizability of each drawing as a function the both the main experimental contrasts (task, site, and age) as well as several effort covariates and individual's estimates tracing abilities.")
```

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & 1.34 & 0.15 & 8.68 & 0.000 \\ 
  Task & -0.04 & 0.26 & -0.16 & 0.873 \\ 
  Age & 0.97 & 0.14 & 7.05 & 0.000 \\ 
  Site & 0.80 & 0.16 & 5.04 & 0.000 \\ 
  Est. tracing score & 0.35 & 0.07 & 4.82 & 0.000 \\ 
  Avg. intensity & 0.06 & 0.02 & 2.47 & 0.013 \\ 
  Draw duration & -0.25 & 0.03 & -7.58 & 0.000 \\ 
  Number of strokes & 0.33 & 0.03 & 10.99 & 0.000 \\ 
  Task*Age & -0.17 & 0.19 & -0.90 & 0.370 \\ 
  Task*Site & -0.22 & 0.16 & -1.38 & 0.168 \\ 
  Age*Site & 0.04 & 0.16 & 0.22 & 0.823 \\ 
  Task*Age*Site & 0.07 & 0.16 & 0.43 & 0.665 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed mode predicting the recognizability of each drawing as a function the both the main experimental contrasts (task, site, and age) as well as several effort covariates and individual's estimates tracing abilities.} 
\end{table}

```{r get-descriptives-across-age}
draw_duration <- d %>%
  group_by(unique_ids,condition,age_numeric, site) %>%
  summarize(avg_draw_duration = mean(draw_duration)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_draw_duration")

num_strokes <- d %>%
  group_by(unique_ids,condition,age_numeric,site) %>%
  summarize(avg_num_strokes = mean(num_strokes)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_num_strokes")

avg_intensity <- d %>%
  group_by(unique_ids,condition,age_numeric,site) %>%
  summarize(avg_intensity = mean(mean_intensity)) %>%
  group_by(age_numeric,condition, site) %>%
  multi_boot_standard(col = "avg_intensity")
```

```{r plot-effort, include=FALSE}
## Make compiled plot of descriptives
base_size_chosen=12 # size of text in plots
smooth_alpha=.2

p1=ggplot(draw_duration, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Draw duration (s)') +
  scale_color_viridis(name="", discrete=TRUE, option='D', begin=.2, end=.8) +
  theme(legend.position = "bottom") + 
  ylim(0,30) +
  facet_grid(~site)

p2=ggplot(avg_intensity, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Ink used (mean intensity)') +
  scale_color_viridis(name="", discrete=TRUE, option='D', begin=.2, end=.8) +
  theme(legend.position = "bottom") + 
  ylim(.02,.08) +
  facet_grid(~site)

p3=ggplot(num_strokes, aes(age_numeric,mean, color=condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = 0.5)) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Number of strokes') +
  theme(legend.position = "bottom") +
  scale_color_viridis(name="", discrete=TRUE, option='D', begin=.2, end=.8) +
  ylim(0,15) +
  facet_grid(~site)
```

```{r effort-covariates, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Effort covariates measured during the drawing task -- amount of time spent drawing, amount of 'ink' used, and number of strokes used -- as function of the age of the child who completed each drawing, the geographical site they were tested at (San Jose vs. Beijing), and the type of drawing task they completed. Error bars show bootstrapped 95 percent cofnidence intervals."}
cowplot::plot_grid(p1,p2,p3,nrow=1)
```

```{r}
tracing_scores_indiv <- d %>%
  ungroup() %>%
  distinct(unique_ids, avg_tracing_score, site, age_numeric)

tracing_scores <- tracing_scores_indiv %>%
  group_by(age_numeric, site) %>%
  multi_boot_standard(col = 'avg_tracing_score')
```

```{r tracing, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Average tracing scores across age and site; each dot represents an average tracing score obtained for each participant and are slightly jittered. Error bars represent bootstrapped 95 percent confidence intervals."} 
ggplot(tracing_scores, aes(x=age_numeric,y=mean, col=site)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_jitter(data = tracing_scores_indiv, aes(x=age_numeric, y=avg_tracing_score), width=.2, height=.01, alpha=.5) +
  theme_few(base_size = base_size_chosen) +
  labs(x='Age', y='Tracing score') +
  theme(legend.position = "none") +
  geom_smooth(col='grey', span = 10,alpha=smooth_alpha)  +
  facet_grid(~site) +
  ylim(-.1,4.1) +
  scale_color_viridis(discrete=TRUE, name="", option='B', begin=0, end=.4)
```

```{r example-drawings, fig.env = "figure", out.width="100%", fig.pos = "H", fig.label="exampledetections", fig.cap = "Randomly sampled, highly recognized drawings for each task, category, and geographical site made by 6-year-old children."}
knitr::include_graphics("figs/example_drawings.png")
```

```{r}
item_effects_not_spread <- d %>%
  group_by(unique_ids,age_numeric,category, site) %>%
  summarize(avg_sub_cor = mean(prop_correct)) %>%
  group_by(age_numeric,category, site) %>%
  multi_boot_standard(col = "avg_sub_cor")
```

```{r item-effects, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=8, fig.cap = "Developmental trajectory of drawing recognizability for each category and for each geographical site. Error bars represent 95 percent bootstrapped confidence intervals."} 
ggplot(item_effects_not_spread, aes(x=as.numeric(age_numeric), y=mean, color=site)) +
  geom_point(alpha=.8) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2), alpha=.8) +
  theme_few(base_size = base_size_chosen) + 
  ylim(0,1) + 
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  ylab('Proportion drawings recognized') +
  xlab('Age (in years)') +
  geom_abline(color = 'grey', intercept=0) +
  geom_smooth(span=20, alpha=.1) +
  facet_wrap(~category) +
  theme(legend.position = 'bottom') +
  scale_color_viridis(discrete=TRUE, name="", option='B', begin=0, end=.4)
```

# General Discussion
Here we examined the developmental trajectory of the recognizability of children's drawings across two different drawing tasks and geographical locations. Overall, we found a consistent developmental trajectory from 4-9 years of age, replicating previous work using observational datasets [@long2021parallel]. We found that asking children to draw categories when provided with a verbal cue ("Can you draw a [cat]?") vs. a picture cue ("Can you draw this [cat] as it looks in the picture?") had little effect on the recognizability of the drawings that children produced. While older children may have integrated specific visual information that was present in the photographs, in keeping with prior work [@bremmer1984prior], this did not seem to change how recognizable their drawings were to outside observers. Thus, these results suggest that younger children's drawings are unlikely to be constrained by their ability to recall specific, canonical exemplars of object categories when they are trying to draw them.

Instead, we found that the recognizability of children's drawings differed between our two different geographical sites -- Beijing, China and San Jose, USA  Why might this be the case?  Our initial hypothesis was that children who had daily practice with a non-orthographic, complex writing system would exhibit greater dexterity in our tracing assessment task -- and that this dexterity could be a factor in how recognizable their drawings were.  However, children across the two sites showed equivalent shape tracing abilities, despite the fact that children's individual tracing scores were strong predictors of the recognizability of the drawings that they produced. A second possibility is that children from preschools near Beijing were simply more engaged with the task than children recruited at a children's museum outside San Jose, USA: A more academic context could have created an environment in which children felt particularly incentivized to spend more effort on their drawings. Nonetheless, our measures of drawing's effort (e.g., amount of time spent drawing, number of strokes made) did not account for differences in recognizability across sites. Furthermore, we saw systematic differences in the object categories that were more or less recognizable across sites, which is not well-explained by this account. Thus, this finding both confirms the idea that children's visuomotor abilities strongly influence the recognizability of the drawings they produce, yet further suggests that it is far from the only factor that does so.

While many have sought to examine variation across different cultural contexts in children's drawings [@winner1989can] or cognitive abilities that may be relevant to children's drawing abilities -- including object recognition [@kuwabara2016cultural] -- we note that this is there is substantial variation within different geographical locations and countries that often goes unmeasured [@amir2020cross]. Indeed, this unmeasured variation within sites may lead to the low replicability of some cross-cultural findings [@carstensen2021investigating]. For example, some work has suggested that the distinction between wheat vs. rice farming traditions explains  variation in cultural practices within China [@talhelm2014large]. We thus caution against a strong interpretation of the observed site differences, as we believe that drawings from children within the same country -- or even area code --  may differ in the very same ways children's drawings differed in this dataset. For example, a final possibility is that the children we recruited from Beijing vs. San Jose could differ widely in the amount of time they spending practicing drawing. If so, then children who spend more time drawing may simply have more experience communicating about the diagnostic features of different object categories.

Of course, these results of course do not preclude the possibility that children's drawings differed across these two drawing tasks or geographical sites in other interesting ways beyond their recognizability. Some work has shown that young children across different cultural contexts tend to produce different tadpole drawings of varying heights and with different amounts of detail [@gernhardt2015cultural] and place a horizon line in systematically different locations when drawing scenes [@senzaki2014holistic]. Thus, finer-grained measurements of these drawings could reveal differences -- with respect to the exact visual features that children chose to include, the strategies they employed while drawing, or the order in which they chose to draw certain elements of these categories.  We have thus made this dataset and the resulting code publicly available for these exploratory analyses.

However, we think a natural question for future work concerns the relationship between the degree to which  difference kinds of developmental experience drawing or learning about visual concepts changes how children draw and recognize them. In a U.S. based sample, prior work found that older children not only produce more recognizable drawings of object categories but tend to be better skilled at recognizing other children's drawings [@long2021parallel], hinting at a relationship between production and recognition. Thus, if children in different environmental contexts tend to spend more time practicing drawing, these children may also be more skilled at recognizing drawings of object categories and perhaps explicitly identifying the distinctive visual features of these categories.

Finally, we note that while this work moves beyond investigating the drawings of children at a single geographical location, these children are still in many ways non-representative of the general population. For example, children in more rural populations may spend considerably less time both consuming and producing drawings of object categories. Thus, these developmental trajectories could thus vary in several ways that are not captured by the populations represented in the the present dataset.

Nonetheless, these results take a first step towards characterizing the consistency in the developmental trajectory of children's drawings.  We propose that the systematic collection of digital drawings from children with a wide variety of experiences will help uncover the myriad factors that drive the developmental changes in children's drawings -- and in turn help us understand how different cognitive systems coordinate to produce this complex behavior.

# Acknowledgements  
[BLINDED]
<!-- We thank Yi Feng and Megan Merrick for assistance with data collection, with whom this project would have not been possible. We also gratefully acknowledge the San Jose Children's Discovery Museum for their collaboration. We are also thankful to all of the parents at XX and XX preschool whose children participated. We also thank the members of the Stanford Language and Cognition lab for their feedback. This work was funded by an NSF SPRF-FR Grant \#1714726 to BLL and a Jacobs Foundation Fellowship to MCF.  -->

\newpage

```{r appendix, include = FALSE}
render_appendix("appendix.Rmd")
```

# References
```{r create_r-references}
r_refs(file = "references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\endgroup