### 2022-07-21 analysis/step1_get_drawings/preprocess_sketches.Rmd
reorg get drawings folder to add scripts to get svg strokes

---
title: "preprocess_sketches CDM"
author: "Bria Long"
date: "9/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(assertthat)
library(knitr)
library(here)
library(lubridate)
```

```{r}
drawing_path = here::here('data/CDM/drawings/full_set/')
filtered_dataset_path =  here::here('data/CDM/drawings/object_drawings')
dir.create(filtered_dataset_path)

compiled_dataset_path = here::here('data/compiled/drawings/object_drawings')
# dir.create(filtered_dataset_path)
dir.create(compiled_dataset_path)
```

```{r}
all_drawings = list.files(path=drawing_path, pattern = ".png$", recursive = TRUE) %>%
    as_tibble()  %>%
    rename(fullPath = value)

practice_categories = c('triangle','this shape','this square','rectangle')

drawing_info = all_drawings %>%
    mutate(fileNameShort = str_split_fixed(fullPath,"/",2)[,2]) %>%
    mutate(condition = str_split_fixed(fileNameShort,"_",2)[,1]) %>%
    mutate(category = str_split_fixed(fileNameShort,"_",3)[,2]) %>%
    mutate(id = str_split_fixed(fileNameShort,"_",4)[,4]) %>%
   select(fileNameShort, fullPath, category, condition) %>%
  filter(!category %in% practice_categories)

```

# copy all drawings and rename
```{r}
file.copy(fs::path(drawing_path, drawing_info$fullPath), fs::path(filtered_dataset_path, drawing_info$fileNameShort))

file.copy(fs::path(drawing_path, drawing_info$fullPath), fs::path(compiled_dataset_path, drawing_info$fileNameShort))

```