### 2022-07-21 analysis/step2_clean_metadata/merge_metadata_and_svgs.Rmd

---
title: "merge_metadata_with_svgs"
author: "Bria Long"
date: '2022-05-25'
output: html_document
---

```{r, libraries}
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(langcog)
library(stringr)
library(here)
theme_set(theme_few())
```

```{r}
meta = read_csv(here::here('data/compiled/metadata/final_merged_metadata.csv'))
```
```{r}
practice_categories = c('triangle','this shape','this square','rectangle')
```

# And now merge with stroke data and resave cleaned versions
```{r}
cdm <- read_csv(here::here('data/CDM/strokes_raw/CDM_photodraw_e2_svg_output2022.csv'))
thu <- read_csv(here::here('data/THU/strokes_raw/THU_photodraw_e2_svg_output2022.csv'))
```

```{r}
all_svgs <- cdm %>%
  select(-'...1') %>%
  mutate(age = as.double(str_split_fixed(age,'age',2)[,2])) %>%
  full_join(thu %>% select(-'...1')) %>%
  select(-filename) %>%
  rename(age_numeric = age)
```

```{r}
missing_subs <- meta %>%
  filter(subID %in% all_svgs$subID) %>%
  distinct(subID)
```

```{r}
all_svgs_with_meta <- all_svgs %>%
  left_join(meta %>% select(-age))
```