### 2022-07-21 analysis/step2_clean_metadata/merge_svg_datasets.Rmd

---
title: "Examine_photodraw_classifications"
author: "Bria Long"
date: "4/2021"
output: html_document
---

```{r, libraries}
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(langcog)
library(stringr)
library(here)
theme_set(theme_few())
```

```{r}
cdm_strokes <- read_csv(here::here('data/CDM/strokes_raw/CDM_photodraw_e2_svg_output2022.csv')) %>%
  select(-filename, -`...1`) %>%
  mutate(session_id = str_split_fixed(session_id, 'CDM_', 2)[,2]) %>%
  mutate(unique_ids = session_id) %>%
  select(-session_id)
```

## Get THU strokes wtih cleaned subid as column
```{r}
thu_joined <- read_csv(here::here('data/THU/strokes_raw/THU_photodraw_e2_svg_output_May2022.csv')) %>%
  select(-filename, -`...1`, -subID) %>%
  rename(unique_ids = subID_cleaned)

length(unique(thu_joined$unique_ids))

```

### Weren't enough subs -- error in subid reading..
Discovered that the subid raw column had spaces that had gotten edited back out, so went through and rewrote stsrokes for those subs
```{r}
# issues <-read_csv(here::here('data/THU/demographics/final_subject_info.csv')) %>%
#   filter(!ID_Database %in% thu_strokes$subID)

```

```{r}
# what<-read_csv(here::here('data/THU/strokes_raw/THU_photodraw_e2_svg_output_May2022.csv'))
#   # filter(is.na(subID))
```

## Import extra strokes
```{r}
all_thu<-read_csv(here::here('data/THU/strokes_raw/THU_photodraw_e2_svg_output_June2022_spaces_added.csv')) %>%
  select(-filename, -`...1`, -subID) %>%
  rename(unique_ids = subID_cleaned) %>%
  full_join(thu_joined)
```
## check subjects
```{r}
length(unique(all_thu$unique_ids))
```

## Join processed thu and cdm data
```{r}
all_strokes <- all_thu %>%
  full_join(cdm_strokes %>% mutate(age = as.double(str_split_fixed(age,'age',2)[,2]))) %>%
  select(-subID) %>%
  filter(! category %in% c('this square','this shape','rectangle','triangle')) %>%
  # P_car_sketch_age5_IPAD3_THU5M7.png
  mutate(filename = paste0(condition,'_',category,'_sketch_age',floor(age),'_', unique_ids,'.png'))
```

## Save it out
```{r}
write_csv(all_strokes,here::here('data/compiled/strokes_preprocessed/all_strokes.csv'))
```

### Compare to metadata

```{r}
# load and clean up "category" for tracing trials
all_meta <- read.csv(here::here('data/compiled/metadata/final_merged_metadata.csv')) %>%
    mutate(category = as.character(category)) %>%
    mutate(category = case_when(category == 'this square' ~ 'square',
                                category == 'this shape' ~ 'shape',                                
                                TRUE ~ as.character(category)))

# subIDs are unique identifier in recognition data for THU, session_ids in CDM data
meta_thu <- all_meta %>%
  filter(site=='THU') %>%
  mutate(unique_ids = subID) %>%
  mutate(draw_duration = draw_duration / 1000) # in ms for thu, whoops

meta_cdm <- all_meta %>%
  filter(site=='CDM') %>%
  mutate(unique_ids = session_id)

## unique IDs are now the identifier which will be used to join with machine/human recognition data
all_meta_cleaned <- meta_cdm %>%
  full_join(meta_thu) 
```

```{r}
unique_ids_strokes = unique(all_strokes$unique_ids)
unique_ids_meta = unique(all_meta_cleaned$unique_ids)
sort(unique_ids_strokes) == sort(unique_ids_meta)
```

```{r}
# Make sure we have all the right filenames
meta_filenames= sort(unique(all_meta_cleaned$filename))
strokes_filenames= sort(unique(all_strokes$filename))
length(strokes_filenames)
length(meta_filenames)

missing <- all_meta_cleaned %>%
  filter(!category %in% c('square','shape','rectangle','triangle')) %>% # exclude shapes
  mutate(filename_check = paste0(condition,'_',category,'_sketch_age',floor(age_numeric),'_', unique_ids,'.png')) %>%
  filter(!filename_check %in% strokes_filenames)

# Draiwngs from the 2 10-year-olds -- have 'age9' so don't get the exact filename match with variables

```

```{r}
# missing_subs = unique(missing$subID)
# 
# missing_subs_demo <-read_csv(here::here('data/THU/demographics/final_subject_info.csv')) %>%
#   filter(ID_Database %in% missing_subs)

```