### 2022-07-21 analysis/step6_main_analyses/clip_classifications.Rmd
scripts for getting devphotodraw clip features

---
title: "clip_classifications"
author: "Bria Long"
date: '2022-05-17'
output: html_document
---

## Load libraries
```{r}
library(tidyverse)
library(here)
library(assertthat)
library(langcog)
library(ggthemes)
library(knitr)
library(lmerTest)
library(lme4)

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## Load data
### Compiled classification data
```{r load-classifications}
# classification_data <- read.csv(here::here('data/compiled/compiled_classifications/Classification_Outputs2760.csv')) %>%
#   as_tibble() %>%
#   mutate(age_numeric = age) %>%
#   mutate(age = paste('age',age,sep="")) %>%
#   mutate(age = as.factor(age)) %>%
#   mutate(category = target_label) %>% 
#   dplyr::select(-X, -X.1, -index) %>%
#   mutate(site = case_when( is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "THU", 
#                            !is.na(str_locate(session_id,'photodraw_e2')[,1]) ~ "CDM")) %>%
#   mutate(site = as.factor(site)) 
```

### Load metadata
```{r}
# load and clean up "category" for tracing trials
all_meta <- read.csv(here::here('data/compiled/metadata/final_merged_metadata.csv')) %>%
    mutate(category = as.character(category)) %>%
    mutate(category = case_when(category == 'this square' ~ 'square',
                                category == 'this shape' ~ 'shape',                                
                                TRUE ~ as.character(category)))

# subIDs are unique identifier in recognition data for THU, session_ids in CDM data
meta_thu <- all_meta %>%
  filter(site=='THU') %>%
  mutate(unique_ids = subID) %>%
  mutate(draw_duration = draw_duration / 1000) # in ms for thu, whoops

meta_cdm <- all_meta %>%
  filter(site=='CDM') %>%
  mutate(unique_ids = session_id)

## unique IDs are now the identifier which will be used to join with machine/human recognition data
all_meta_cleaned <- meta_cdm %>%
  full_join(meta_thu)

```

```{r}
all_meta_cleaned <- all_meta_cleaned %>%
  select(-age) %>%
  rename(age = age_numeric)
```

### Load clip classifications
```{R}
clip_features <- read.csv(here::here('data/compiled/clip_classifications/CLIP_FEATURES_kid_photodraw_compiled_drawing_cue.csv'), header=TRUE) %>%
  as_tibble()

clip_meta <- read.csv(here::here('data/compiled/clip_classifications/CLIP_METADATA_batch32_kid.csv')) %>%
  as_tibble() %>%
  rename(category = label, session_id = session)

all_clip = clip_features %>%
  right_join(clip_meta, by=c('X')) 
```

#### Make long form clip data
```{r}
clip_long <- all_clip %>%
  rename(kid_age = age, site = condition, unique_ids = session_id) %>%
  select(-X) %>%
  pivot_longer(cols = cat:tree, names_to  = "guessed_category", values_to ="prob")  %>%
  mutate(guessed_category = factor(guessed_category)) %>%
  mutate(category = factor(category))
```

```{r}
# test_image = 'P_airplane_sketch_age8_IPAD2_THU8F12.png'
# test_session_id = 'IPAD2_THU8F12'

# test_image_clip <- all_clip %>%
  # filter(session_id == test_session_id & category=='airplane')
```

```{r}
clip_by_image <- clip_long %>%
  group_by(kid_age, unique_ids, category) %>%
  summarize(max_prob = max(prob), clip_category = guessed_category[prob == max_prob]) %>%
  mutate(correct_or_not = (category == clip_category)) %>%
  left_join(all_meta_cleaned)
```

```{r}
count_images <- clip_by_image %>% 
  group_by(kid_age, category) %>%
  summarize(count = n())

check <- clip_by_image %>% 
  filter(is.na(condition))

sum(count_images$count) # sanity
```

```{r}
clip_target_category_prob_by_age_by_category <- clip_long %>%
  filter(category == guessed_category) %>%
  group_by(category, kid_age) %>%
  multi_boot_standard(col = 'prob')

target_category_prob_by_age <- clip_target_category_prob_by_age_by_category %>%
  group_by(kid_age) %>%
  multi_boot_standard(col = 'mean')
```

```{r}
clip_correct_category_by_age <- clip_by_image %>%
   mutate(age_numeric = as.double(kid_age)) %>%
  mutate(age_numeric = case_when(age_numeric == 10 ~ 9, # merge 2 10-year-olds into THU data for 9-year-olds
                                 TRUE ~ age_numeric)) %>%
  group_by(category, age_numeric) %>%
  multi_boot_standard(col = 'correct_or_not') 
```

```{r}
ggplot(clip_correct_category_by_age, aes(x=age_numeric, y=mean, col=age_numeric)) +
  geom_pointrange(aes(y=mean, ymin = ci_lower, ymax = ci_upper), alpha=.3)+
  geom_point(alpha=.6) +
  # geom_line(data = clip_correct_category_by_age, alpha=.4, aes(group=category))+
  geom_smooth(alpha=.01, color='grey', span=2) +
  labs(x='Age of child drawing (yrs)', y='CLIP classification accuracy') +
  theme_few() + 
  scale_x_continuous(breaks = seq(2,10,1)) + 
  scale_color_viridis(option="D", breaks=seq(2,10,1)) +
  theme(legend.position= 'none') +
  facet_wrap(~category)

```

### Load human classification data
```{r}
humans <- read.csv(here::here('data/compiled/recognition_ratings/compiled_human_recognition.csv'))

humans_merged <- humans %>%
  group_by(unique_ids, category, image_name_short) %>% # group by each drawing of each category
  summarize(num_correct = mean(correct_or_not)) %>%
  left_join(clip_by_image, by=c('unique_ids','category')) %>% # join with classification data 
  filter(!is.na(age))
```

```{r}
ggplot(humans_merged, aes(x=correct_or_not, y=num_correct)) +
  geom_jitter(alpha=.2, width=.1, height=.1) +
  geom_smooth(method='lm') +
  ylab('human recognition') +
  xlab('clip model recognition')+
  theme_few()
```

```{r}
by_category <- humans_merged %>%
  group_by(category) %>%
  summarize(humans_pc = mean(num_correct), model_pc = mean(correct_or_not, na.rm=TRUE))

ggplot(by_category, aes(x=model_pc, y=humans_pc, label=category)) +
  geom_point() +
  geom_smooth(method='lm') +
  ylab('human recognition') +
  xlab('model recognition') +
  theme_few() +
  ggrepel::geom_label_repel()
```

### Correlates humans vs vgg classifications 
```{r}
# cor.test(humans_merged$num_correct, humans_merged$target_label_prob)
cor.test(humans_merged$num_correct, as.numeric(as.logical(humans_merged$correct_or_not)))
```

```{r}

category_by_age_by_cond <- humans_merged %>%
  group_by(category, age, condition) %>%
  summarize(humans_pc = mean(num_correct), model_pc = mean(correct_or_not, na.rm=TRUE))  %>%
  mutate(diff = (model_pc - humans_pc))

```

```{r}
ggplot(category_by_age_by_cond, aes(x=model_pc, y=humans_pc,color=age, label=category))+
  geom_point() +
  ylim(0,1) +
  xlim(0,1) +
  theme_few() +
  theme(aspect.ratio = 1) +
  facet_wrap(~category) +
  scale_color_viridis_c() +
  geom_abline(xintercept=0, color='grey')
  ggrepel::geom_label_repel(data = category_by_age_by_cond %>% filter(diff>.1), label.size=.1, max.overlaps=20) 
```

```{r}
cor_by_age_by_site_by_cond <- humans_merged %>%  
  group_by(unique_ids,category,condition,age, site) %>%
  summarize(avg_cor = mean(correct_or_not)) %>%
  group_by(age,condition, site) %>%
  multi_boot_standard(col = "avg_cor") %>%
  mutate(who = "model")

cor_by_age_by_site_by_cond_humans <- humans_merged %>%  
  group_by(unique_ids,category,condition,age, site) %>%
  summarize(avg_cor = mean(num_correct)) %>%
  group_by(age,condition, site) %>%
  multi_boot_standard(col = "avg_cor") %>%
  mutate(who = 'humans')

humans_and_models <- cor_by_age_by_site_by_cond %>%
  full_join(cor_by_age_by_site_by_cond_humans)

```

```{r}
summary(lmer(data=humans_merged, correct_or_not ~ condition*age + (1|unique_ids) + (1|category)))
```

```{r}
summary(lmer(data=humans_merged, num_correct ~ condition*age + (1|unique_ids) + (1|category)))

```

```{r}
base_size_chosen=12; smooth_alpha=.2
ggplot(humans_and_models, aes(age,mean, col = condition)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.2)) +
  theme_few(base_size = base_size_chosen) + 
  labs(x='Age', y='Classification accuracy') +
  ylim(0,1) + 
  geom_smooth(span=10, alpha=.2) +
  geom_hline(yintercept = 1/12, linetype="dashed", color="grey") +
  facet_grid(who~site)
```

## How well does the model do on the images that humans unanimously recognize?
```{r}
easiest_to_recognize <- humans_merged %>%
  filter(num_correct == 1) %>%
  group_by(category) %>%
  summarize(model_correct = mean(correct_or_not), num_images= n()) %>%
  mutate(category = fct_reorder(category, -model_correct))
```

```{r}
clip_to_merge <- clip_by_image %>%
  select(-age) %>%
  rename(age = kid_age) %>% 
  select(age, unique_ids, category, correct_or_not) %>%
  mutate(model = 'clip')

vgg_vs_clip <- classification_data %>%
  select(-age) %>%
  rename(age = age_numeric, unique_ids = session_id) %>%
  mutate(model = 'vgg') %>%
  select(unique_ids, age, category, model, correct_or_not) %>%
  full_join(clip_to_merge)

```

```{r}
vgg_vs_clip_vs_humans <- vgg_vs_clip %>%
  left_join(humans_merged %>% select(unique_ids, category, num_correct) %>% rename(human_correct = num_correct))

```

```{r}
vgg_vs_clip_easy <- vgg_vs_clip_vs_humans %>%
  filter(human_correct == 1) %>%
  group_by(model, category) %>%
  summarize(model_correct = mean(correct_or_not), num_images = n()) %>%
  mutate(category = fct_reorder(category, -model_correct))
```
```{r}
vgg_vs_clip_easy_stats <- vgg_vs_clip_easy %>%
  summarize(avg_model_correct = mean(model_correct), sd_model_correct = sd(model_correct))

```

```{r}
summary(lmer(data = vgg_vs_clip_easy, model_correct ~ model + (1|category)))
```

```{r}
ggplot(vgg_vs_clip_easy, aes(x=category, y=model_correct, color=model)) +
  geom_point(aes(size = num_images)) +
  theme_few() +
  ylab('Model percent correct') +
  ylim(0,1) +
  ggtitle('CLIP vs VGG performance on very recognizable drawings')

```

```{r}

ggplot(easiest_to_recognize, aes(x=category, y=model_correct)) +
  geom_point(aes(size = num_images)) +
  theme_few() +
  ylab('CLIP percent correct') +
  ggtitle('CLIP performance on very recognizable drawings')

```

```{r}
ggplot(vgg_vs_clip_vs_humans, aes(x=correct_or_not, y=human_correct, color=model)) +
  theme_few() +
  geom_smooth(aes(group=model), method='lm') +
  geom_jitter(width=.2, height=.01, alpha=.2) +
  facet_grid(~model) +
  ylab('Average human recognition') +
  xlab('Model recognition') +
  scale_x_continuous(breaks = c(0,1))

```

```{r}
ggplot(vgg_vs_clip_vs_humans, aes(x=correct_or_not, y=human_correct, color=model)) +
  theme_few() +
  geom_smooth(aes(group=model), method='lm') +
  geom_jitter(width=.1, height=.01, alpha=.2) +
  facet_wrap(~category) +
  ylab('Average human recognition') +
  xlab('Model recognition')
```